### 2016-04-12 manuscript_v0.Rmd
Getting Started

---
title: "Testing new Pauly Cessational Growth Function"
author: "Derek H. Ogle"
date: "April 9, 2016"
output: 
  pdf_document: 
    fig_caption: yes
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FSA)
library(FSAsim)

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
vbSO <- vbFuns("Somers2")
```

```{r captions, include=FALSE}

```

```{r maketprFun, echo=FALSE}
################################################################################
## internal function to compute t-prime
## 
################################################################################
iCalc_tpr <- function(t,WP,NGT) {
  ## First NGT stats at WP, use this to shift each age such that
  ##    first NGT would correspond to a value of 1
  tshift <- 1-WP+NGT/2
  tmp.t <- t+tshift
  ## Find fraction of year on this new time scale for each adjusted age
  tmp.t2 <- tmp.t-floor(tmp.t)
  ## Adjust this fraction for no growth
  for (i in 1:length(tmp.t2)) {
    if (tmp.t2[i]<=NGT) tmp.t2[i] <- 0
    else tmp.t2[i] <- tmp.t2[i]-NGT
  }
  tmp.t <- floor(tmp.t)*(1-NGT)+tmp.t2
  ## Shift back to get tprime (why is NGT needed here)
  tpr <- tmp.t-tshift+NGT
}
```

```{r makePaulyFun, echo=FALSE}
################################################################################
## Main Function
##
##   Linf, t0 as usual
##   Kpr = K-prime as defined in Pauly et al. (units are diff than usual K)
##   WP = "Winter Period" (point where growth=0) = ts+0.5 (ts from Pauly et al.)
##   NGT = "No Growth Time" = "fraction of a year where no growth occurs
##
##   tpr = "t-prime" = actual age (t) minus cumulative NGT prior to t
##   Q is as defined in Pauly et al.
##   qt and qtr are intermediate values to make Pauly SC look similar to the
##     Somers and Somers2 models from vbFuns() in FSA
##
##   The final line is basically Equation 4 from Pauly et al.
################################################################################

paulySC <- function(t,Linf,Kpr=NULL,t0=NULL,WP=NULL,NGT=NULL) {
  if (length(Linf)==5) { Kpr <- Linf[[2]]; t0 <- Linf[[3]]
                         WP <- Linf[[4]]; NGT <- Linf[[5]]
                         Linf <- Linf[[1]] }
  tpr <- iCalc_tpr(t,WP,NGT)
  Q <- (2*pi)/(1-NGT)
  qt <- (Kpr/Q)*sin(Q*(tpr-WP+0.5))
  qt0 <- (Kpr/Q)*sin(Q*(t0-WP+0.5))
  Linf*(1-exp(-Kpr*(tpr-t0)-qt+qt0))
}
```

# Introduction

The growth of many animals exhibits seasonal oscillations in response to seasonal changes in temperature, light, and food supplies ().  Several functions have been proposed to model seasonal growth.  The most popular of these functions, from Somers (1988) but carefully reiterated in GBO (), is

\[ L_{t} = L_{\infty}(1-e^{-K(t-t_{0})}-S(t)+S(t_{0}))  \]

with $S(t) = \frac{CK}{2\pi}sin(2\pi(t-t_{s}))$ so that $S(t_{0}) = \frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))$ and where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the model asymptote for average length, $K$ is a measure of the exponential rate of approach to the asymptotic length (Schnute and Fournier 1980), $t_{0}$ is the theoretical time or age (generally negative) at which the average length would be zero, $C$ modulates the amplitude of the growth oscillations and corresponds to the proportion of decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and this function reduces to the typical von Bertalanffy growth function.  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter.

Values of C[1 (or\0)
allow seasonal decreases in average length-at-age,
and might not seem realistic in organisms whose
skeletons largely preclude shrinkage (Pauly et al.
1992), but could result from size-dependent overwinter
mortality.

1. Seasonal Growth
1. Somer's Seasonal Growth Function (SoSVBGF)

```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.  The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.8; C <- 1
Linf <- 30; K <- 0.7; t0 <- -0.05
Scf <- c(Linf,K,t0,C,WP)
t <- seq(0,5,length.out=299)
SL <- vbSO(t,Linf=Scf)
plot(SL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
Scf <- c(Linf,K,t0,C=2,WP)
SL <- vbSO(t,Linf=Scf)
lines(SL~t,lwd=2,lty=2)
Scf <- c(Linf,K,t0,C=0.2,WP)
SL <- vbSO(t,Linf=Scf)
lines(SL~t,lwd=2,lty=3)

WPs <- 0:5+WP
LatWPs <- vbSO(WPs,Linf=Scf)
segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")

legend("topleft",c("C=1","C=2","C=0.2"),lwd=2,lty=1:3,bty="n",inset=-0.02)
```

1. Pauly's Seasonal Growth Function (PaSVBGF)
    1. Define terms
    1. Explain t-prime
    1. Explain K-prime

```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.  The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.8; NGT <- 0.4
Linf <- 30; Kpr <- 0.7; t0 <- -0.05
Pcf <- c(Linf,Kpr,t0,WP,NGT)
t <- seq(0,5,length.out=299)
PL <- paulySC(t,Linf=Pcf)
plot(PL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
WPs <- 0:5+WP
LatWPs <- paulySC(WPs,Linf=Pcf)
segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")
segments(WPs+NGT,0,WPs+NGT,LatWPs,lty=3,lwd=2,col="blue")
legend("topleft",c("WP=0.8","NGT=0.4"),bty="n",inset=-0.02,text.col=c("red","blue"))
```

# Methods
## Simulated Data

```{r echo=FALSE}
WP <- 0.8
NGT <- 0.5
gper <- c(WP+NGT/2-1,WP-NGT/2)*365
## Set some growth parameters
Linf <- 30; K <- 0.3; t0 <- 0
## Actually generate the data (with little variability)
d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.01,SD=0.01,paramCV=0.01,
                growth.per=gper,sample.per=c(1,365),lendigs=1)
```

## R Implementation

# Model Fitting
## Simulated Data

```{r echo=FALSE, message=FALSE}
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
## Try fitting the PaulySC model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.8,NGT=0.1)
fitP1 <- nls(lenCap~paulySC(ageFracY,Linf,K,t0,WP,NGT),data=d1,
             start=sv,lower=lwrbnd,upper=uprbnd,algorithm="port",control=list(maxiter=100))
## Make a diagnostic plot
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5))
curve(paulySC(x,coef(fitP1)),from=1,to=6,col="red",lwd=3,add=TRUE)
cfs <- coef(fitP1)
WPs <- 1:5+cfs[["WP"]]
LatWPs <- paulySC(WPs,Linf=cfs)
segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col="red")
SNGTs <- WPs-cfs[["NGT"]]/2
ENGTs <- WPs+cfs[["NGT"]]/2
segments(SNGTs,LatWPs-1,SNGTs,LatWPs+1,lty=3,lwd=2,col="blue")
segments(ENGTs,LatWPs-1,ENGTs,LatWPs+1,lty=3,lwd=2,col="blue")
legend("topleft",paste(c("WP","NGT"),c(WP,NGT),sep="="),
       bty="n",inset=-0.02,text.col=c("red","blue"))

cbind(Est=coef(fitP1),confint(fitP1))
```
```{r echo=FALSE}
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5),xlim=c(1.7,1.9),ylim=c(13.425,13.43))
curve(paulySC(x,coef(fitP1)),from=1,to=6,col="red",lwd=3,add=TRUE,n=1001)
segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col="red")
segments(SNGTs,LatWPs-1,SNGTs,LatWPs+1,lty=3,lwd=2,col="blue")
segments(ENGTs,LatWPs-1,ENGTs,LatWPs+1,lty=3,lwd=2,col="blue")

```

```{r echo=FALSE, message=FALSE}
## Try fitting the Somers model
vbSO <- vbFuns("Somers2")
fitSO1 <- nls(lenCap~vbSO(ageFracY,Linf,K,t0,C,WP),data=d1,
               start=list(Linf=32,K=0.32,t0=0,C=0.8,WP=0.8),
               lower=c(Linf=0,K=0,t0=-Inf,C=-1,WP=0),
               upper=c(Linf=Inf,K=Inf,t0=Inf,C=2,WP=1),
               algorithm="port",control=list(maxiter=100),trace=FALSE)
plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5))
curve(vbSO(x,coef(fitSO1)),from=1,to=6,col="red",lwd=2,add=TRUE)
cbind(coef(fitSO1),confint(fitSO1))
```

## Real Data

## Appendix

```{r}
<<maketprFun>>
```

```{r}
<<makePaulyFun>>
```

### 2016-04-12 zzzOld/zzzFirstTry/seasonalCessation.R

@@ -1,0 +1,157 @@
+################################################################################
+## This is an attempt at a function to create the Pauly et al. (1992)
+## seasonal cessation Von B growth model
+##
+##   Linf, K, t0 as usual
+##   WP = "Winter Period" (point where growth=0) = ts+0.5 (ts from Pauly et al.)
+##   NGT = "No Growth Time" = "fraction of a year where no growth occurs
+##
+##   tpr = "t-prime" = actual age (t) minus cumulative NGT prior to t
+##   Q is as defined in Pauly et al.
+##   qt and qtr are intermediate values to make Pauly SC look similar to the
+##     Somers and Somers2 models from vbFuns() in FSA
+##
+##   The final line is basically Equation 4 from Pauly et al.
+################################################################################
+paulySC <- function(t,Linf,K,t0,WP,NGT) {
+  ## Shift time so that first NGT is at 1
+  # NGT starts at WP
+  shift <- 1-WP
+  tmp.t <- t+shift
+  ## Find fraction of year on this new time scale
+  tmp.t2 <- tmp.t-floor(tmp.t)
+  ## Adjust this fraction for no growth
+  for (i in 1:length(tmp.t2)) {
+    if (tmp.t2[i]<=NGT) tmp.t2[i] <- 0
+    else tmp.t2[i] <- tmp.t2[i]-NGT
+  }
+  tmp.t <- floor(tmp.t)*(1-NGT)+tmp.t2
+  ## Shift back to get tprime
+  tpr <- tmp.t-shift+NGT/2  # why is NGT/2 needed here
+  ## return model values
+  Q <- (2*pi)/(1-NGT)
+  qt <- (K/Q)*sin(Q*(tpr-WP+0.5))
+  qt0 <- (K/Q)*sin(Q*(t0-WP+0.5))
+  Linf*(1-exp(-K*(tpr-t0)-qt+qt0))
+}
+
+
+################################################################################
+## Some Checks?
+################################################################################
+## Does this basically reproduce Figure 1.4 in Magnifico ... IT DOES NOT
+par(xaxs="i",yaxs="i")
+t <- seq(-1,5,0.01)
[…]

### 2016-04-13 manuscript_v0.Rmd
Initial work on manuscript

@@
 output:
+  html_document: 
+    fig_height: 3.5
+    fig_width: 4
   pdf_document:
     fig_caption: yes
+csl: american-fisheries-society.csl
+bibliography: SeasonalGrowth.bib
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+set.seed(34738423)  # for reproducibility
 library(FSA)
@@
 eqncaps <- captioner(prefix="Equation")
+eqncaps("VBMain")
+eqncaps("SomersMod")
+eqncaps("PaulyMod")
+eqncaps("PaulyMod2")
+
 vbSO <- vbFuns("Somers2")
@@
 Several functions have been proposed to model seasonal growth.
-The most popular of these functions, from Somers (1988) but carefully reiterated in GBO (), is
+The most popular of these functions, from XXX and @Somers1988 and carefully reiterated in @GarciaBerthouetal2012, is
 
-\[ L_{t} = L_{\infty}(1-e^{-K(t-t_{0})}-S(t)+S(t_{0}))  \]
+$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("VBMain",display="num"),")")`} $$
 
-with $S(t) = \frac{CK}{2\pi}sin(2\pi(t-t_{s}))$ so that $S(t_{0}) = \frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))$ and where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the model asymptote for average length, $K$ is a measure of the exponential rate of approach to the asymptotic length (Schnute and Fournier 1980), $t_{0}$ is the theoretical time or age (generally negative) at which the average length would be zero, $C$ modulates the amplitude of the growth oscillations and corresponds to the proportion of decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).
-If $C$=0, then there is no seasonal oscillation and this function reduces to the typical von Bertalanffy growth function.
-If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter.
+with
 
-Values of C[1 (or\0)
-allow seasonal decreases in average length-at-age,
-and might not seem realistic in organisms whose
-skeletons largely preclude shrinkage (Pauly et al.
-1992), but could result from size-dependent overwinter
-mortality.
+$$ q=K(t-t_{0})+\frac{CK}{2\pi}sin(2\pi(t-t_{s}))-\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s})) \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`} $$
 
-1. Seasonal Growth
-1. Somer's Seasonal Growth Function (SoSVBGF)
+where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the model asymptote for average length, $K$ is a measure of the exponential rate of approach to the asymptotic length [@SchnuteFournier1980], $t_{0}$ is the theoretical time or age (generally negative) at which the average length would be zero, $C$ modulates the amplitude of the growth oscillations and corresponds to the proportion of decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).
+If $C$=0, then there is no seasonal oscillation and this function reduces to the typical von Bertalanffy growth function (`r figcaps("SomersExPlot",display="cite")`).
+If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).
 
-```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.
-The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
+\
+
+```{r echo=FALSE, results="hide"}
 par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
-WP <- 0.8; C <- 1
-Linf <- 30; K <- 0.7; t0 <- -0.05
-Scf <- c(Linf,K,t0,C,WP)
-t <- seq(0,5,length.out=299)
-SL <- vbSO(t,Linf=Scf)
-plot(SL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
-Scf <- c(Linf,K,t0,C=2,WP)
-SL <- vbSO(t,Linf=Scf)
-lines(SL~t,lwd=2,lty=2)
-Scf <- c(Linf,K,t0,C=0.2,WP)
-SL <- vbSO(t,Linf=Scf)
-lines(SL~t,lwd=2,lty=3)
+WP <- 0.8; Linf <- 30; K <- 0.3; t0 <- -0.05
+C <- c(0,0.5,1,2)
+clr <- col2rgbt(rep("black",4),1/(1:4))
+t <- seq(0,3,length.out=299)
+plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],WP))~t,type="l",lwd=2,col=clr[1],
+     ylim=c(0,20),ylab="Length",xlab="Age (years)",xaxt="n")
+axis(1,0:3)
+lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],WP))~t,lwd=2,col=clr[2])
+lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],WP))~t,lwd=2,col=clr[3])
+lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],WP))~t,lwd=2,col=clr[4])
 
-WPs <- 0:5+WP
-LatWPs <- vbSO(WPs,Linf=Scf)
-segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")
+#WPs <- 0:5+WP
+#LatWPs <- vbSO(WPs,Linf=Scf)
+#segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")
 
-legend("topleft",c("C=1","C=2","C=0.2"),lwd=2,lty=1:3,bty="n",inset=-0.02)
+legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)
+
+figcaps("SomersExPlot",paste0("Example growth trajectory of ",eqncaps('VBMain',display='cite')," using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
 ```
 
-1. Pauly's Seasonal Growth Function (PaSVBGF)
-    1. Define terms
-    1. Explain t-prime
-    1. Explain K-prime
+`r figcaps("SomersExPlot")`
 
-```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.
-The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
+\
+
+Values of $C$&gt;1 (or &lt;0) allow seasonal decreases in average length-at-age (`r figcaps("SomersExPlot",display="cite")`).
+A seasonal decrease in average length-at-age could result from size-dependent overwinter mortality [@GarciaBerthouetal2012], but this growth trajectory is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992].
+
+@Paulyetal1992 introduced a modification of `r eqncaps("SomersMod")` that modeled a true no-growth period and that smoothly transitioned into and out of the no-growth period.
+Specifically, their function is `r eqncaps("VBMain")` with
+
+\
+
+$$ q=K(t^{'}-t_{0})+\frac{K(1-NGT)}{2\pi}sin(\frac{2\pi}{1-NGT}(t^{'}-t_{s}))-\frac{K(1-NGT)}{2\pi}sin(\frac{2\pi}{1-NGT}(t_{0}-t_{s})) \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`} $$
+
+\
+
+\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (in fractions of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  @Paulyetal1992 introduced $K^{'}$ to `r eqncaps("PaulyMod",display="cite")` as the units of $K$ change from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")` because $t^{'}$&lt;$t$ for any $t$.
+We further note here that $K^{'}$&gt;$K$ because annual growth is compressed into a shorter period ($1-NGT$) in `r eqncaps("PaulyMod",display="cite")` when $NGT$&gt;0.
+
+@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at one point), replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the growth period), and replacing $K$ with $K^{'}$.
+Furthermore, their modification can be described geometrically (though not algorithmically) in two steps.
+First, the seasonal growth function combined in Equations `r eqncaps("VBMain",display="num")` and `r eqncaps("SomersMod",display="num")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative no-growth time subtracted (i.e., the $t^{'}$).
+The growth trajectory is then separated at the $WP$ and a segment with a slope of zero that is $NGT$ units long is inserted into the trajectory.
+This forms a trajectory that smoothly transitions into and out of a no-growth period (`r figcaps("PaulyExPlot")).
+
+\
+
+```{r echo=FALSE, results="hide"}
 par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
-WP <- 0.8; NGT <- 0.4
-Linf <- 30; Kpr <- 0.7; t0 <- -0.05
+clrs <- c("black",col2rgbt("black",1/2))
+WP <- 0.8; NGT <- 0.3
+Linf <- 30; Kpr <- 0.4; t0 <- -0.05
 Pcf <- c(Linf,Kpr,t0,WP,NGT)
-t <- seq(0,5,length.out=299)
+t <- seq(0,3,length.out=299)
 PL <- paulySC(t,Linf=Pcf)
-plot(PL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
-WPs <- 0:5+WP
+plot(PL~t,type="l",lwd=2,ylim=c(0,20),ylab="Length",xlab="Age (years)")
+WPs <- 0:3+WP
 LatWPs <- paulySC(WPs,Linf=Pcf)
-segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")
-segments(WPs+NGT,0,WPs+NGT,LatWPs,lty=3,lwd=2,col="blue")
-legend("topleft",c("WP=0.8","NGT=0.4"),bty="n",inset=-0.02,text.col=c("red","blue"))
+segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col=clrs[1])
+segments(WPs-NGT/2,LatWPs-1,WPs-NGT/2,LatWPs+1,lty=3,lwd=2,col=clrs[2])
+segments(WPs+NGT/2,LatWPs-1,WPs+NGT/2,LatWPs+1,lty=3,lwd=2,col=clrs[2])
+figcaps("PaulyExPlot",paste0("Example growth trajectory of ",eqncaps('VBMain',display='cite')," using ",eqncaps('PaulyMod',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $WP$=",WP,", and $NGT$=",NGT,".
+Each $WP$ is marked by the longer vertical lines and the beginning and ending of the no-growth times are marked by the paired shorter vertical lines."))
 ```
+
+`r figcaps("PaulyExPlot")`
+
+\
 

### 2016-04-14 manuscript_v0.Rmd
More work on ms

--- 2016-04-12 manuscript_v0.Rmd
@@
   pdf_document:
-    fig_caption: yes
+    fig_height: 3.5
+    fig_width: 4.5
+  html_document:
+    fig_height: 3.5
+    fig_width: 4.5
+csl: american-fisheries-society.csl
+bibliography: SeasonalGrowth.bib
 ---
@@
 eqncaps <- captioner(prefix="Equation")
+eqncaps("vbMain")
+eqncaps("vbMain2")
+eqncaps("SomersMod")
+eqncaps("PaulyMod")
+eqncaps("PaulyMod2")
+
 vbSO <- vbFuns("Somers2")
@@
   ## First NGT stats at WP, use this to shift each age such that
-  ##    first NGT would correspond to a value of 1
-  tshift <- 1-WP+NGT/2
-  tmp.t <- t+tshift
+  ##    first NGT would correspond to a value of 0
+  tshift <- WP-NGT/2
+  tmp.t <- t-tshift
   ## Find fraction of year on this new time scale for each adjusted age
@@
   ## Shift back to get tprime (why is NGT needed here)
-  tpr <- tmp.t-tshift+NGT
+  tpr <- tmp.t+tshift+NGT
 }
@@
   Q <- (2*pi)/(1-NGT)
-  qt <- (Kpr/Q)*sin(Q*(tpr-WP+0.5))
-  qt0 <- (Kpr/Q)*sin(Q*(t0-WP+0.5))
+  qt <- (Kpr/Q)*sin(Q*(tpr-WP-0.5))
+  qt0 <- (Kpr/Q)*sin(Q*(t0-WP-0.5))
   Linf*(1-exp(-Kpr*(tpr-t0)-qt+qt0))
@@
 
-The growth of many animals exhibits seasonal oscillations in response to seasonal changes in temperature, light, and food supplies ().
-Several functions have been proposed to model seasonal growth.
-The most popular of these functions, from Somers (1988) but carefully reiterated in GBO (), is
+The mean length-at-age for many animals (e.g., ) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).
+The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as
 
-\[ L_{t} = L_{\infty}(1-e^{-K(t-t_{0})}-S(t)+S(t_{0}))  \]
+$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("vbMain",display="num"),")")`} $$
 
-with $S(t) = \frac{CK}{2\pi}sin(2\pi(t-t_{s}))$ so that $S(t_{0}) = \frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))$ and where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the model asymptote for average length, $K$ is a measure of the exponential rate of approach to the asymptotic length (Schnute and Fournier 1980), $t_{0}$ is the theoretical time or age (generally negative) at which the average length would be zero, $C$ modulates the amplitude of the growth oscillations and corresponds to the proportion of decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).
-If $C$=0, then there is no seasonal oscillation and this function reduces to the typical von Bertalanffy growth function.
-If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter.
+with
 
-Values of C[1 (or\0)
-allow seasonal decreases in average length-at-age,
-and might not seem realistic in organisms whose
-skeletons largely preclude shrinkage (Pauly et al.
-1992), but could result from size-dependent overwinter
-mortality.
+$$ q=K(t-t_{0}) \quad \quad \text{`r paste0("(",eqncaps("vbMain2",display="num"),")")`} $$
 
-1. Seasonal Growth
-1. Somer's Seasonal Growth Function (SoSVBGF)
+where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the asymptotic mean length, $K$ is a measure of the exponential rate of approach to the asymptote [@SchnuteFournier1980], and $t_{0}$ is the theoretical time or age (generally negative) at which the mean length would be zero.
 
-```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.
-The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
+Many animals exhibits seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply ().
+`r eqncaps("vbMod",display="cite")` of the traditional VBGF has been modified, usually with a sin function, to model these seasonal oscillations in growth.
+The most popular of these modifications is from @HoenigChoudarayHanumara1982 and @Somers1988 (and carefully reiterated in @GarciaBerthouetal2012), and uses
+
+$$
+\begin{aligned}
+q & =K(t-t_{0}) \\
+  & +\frac{CK}{2\pi}sin(2\pi(t-t_{s}))  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`} \\
+  & -\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))
+\end{aligned}  
+$$
+
+where $C$ modulates the amplitude of the growth oscillations and corresponds to the proportional decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).
+If $C$=0, then there is no seasonal oscillation and `r eqncaps("SomersMod",display="cite")` reduces to  `r eqncaps("vbMain2",display="cite")` and the typical VBGF (`r figcaps("SomersExPlot",display="cite")`).
+If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).
+
+\
+
+```{r echo=FALSE, results="hide"}
 par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
-WP <- 0.8; C <- 1
-Linf <- 30; K <- 0.7; t0 <- -0.05
-Scf <- c(Linf,K,t0,C,WP)
-t <- seq(0,5,length.out=299)
-SL <- vbSO(t,Linf=Scf)
-plot(SL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
-Scf <- c(Linf,K,t0,C=2,WP)
-SL <- vbSO(t,Linf=Scf)
-lines(SL~t,lwd=2,lty=2)
-Scf <- c(Linf,K,t0,C=0.2,WP)
-SL <- vbSO(t,Linf=Scf)
-lines(SL~t,lwd=2,lty=3)
+WP <- 0.55; Linf <- 30; K <- 0.3; t0 <- -0.1
+C <- c(0,0.5,1,2)
+clr <- col2rgbt(rep("black",4),1/(1:4))
+t <- seq(0,3,length.out=299)
+plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],WP))~t,type="l",lwd=2,col=clr[1],
+     ylim=c(0,20),ylab="Length",xlab="Age (years)",xaxt="n")
+axis(1,0:3)
+lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],WP))~t,lwd=2,col=clr[2])
+lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],WP))~t,lwd=2,col=clr[3])
+lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],WP))~t,lwd=2,col=clr[4])
 
-WPs <- 0:5+WP
-LatWPs <- vbSO(WPs,Linf=Scf)
-segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")
+legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)
 
-legend("topleft",c("C=1","C=2","C=0.2"),lwd=2,lty=1:3,bty="n",inset=-0.02)
+figcaps("SomersExPlot",paste0("Example growth trajectory of ",eqncaps('vbMain',display='cite')," using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
 ```
 
-1. Pauly's Seasonal Growth Function (PaSVBGF)
-    1. Define terms
-    1. Explain t-prime
-    1. Explain K-prime
+`r figcaps("SomersExPlot")`
 
-```{r echo=FALSE, fig.width=5, fig.height=4, fig.cap="Example growth trajectory of Equation X using Linf=30, Kpr=0.7, t0=-0.05, WP=0.8, and NGT=0.4.
-The start of each no growth time (i.e, Annual Age+WP) is marked by a red vertical line and the end of each no growth time (i.e., Annual Age+WP+NGT)."}
-par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
-WP <- 0.8; NGT <- 0.4
-Linf <- 30; Kpr <- 0.7; t0 <- -0.05
+\
+
+Values of $C$&gt;1 (or &lt;0) in  `r eqncaps("SomersMod",display="cite")` allow seasonal decreases in mean length-at-age (`r figcaps("SomersExPlot",display="cite")`).
+A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal crease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].
+To address this issue, @Paulyetal1992 modified `r eqncaps("SomersMod")` to include a true seasonal no-growth period with a smooth transition of the modeled mean length-at-age into and out of the no-growth period.
+Specifically, their modification is
+
+\
+
+$$
+\begin{aligned}
+q & =K^{'}(t^{'}-t_{0}) \\
+  & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-t_{s})\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`} \\
+  & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-t_{s})\right)
+\end{aligned}  
+$$
+
+\
+
+\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (in fractions of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 introduced $K^{'}$ to `r eqncaps("PaulyMod",display="cite")` because the units of $K$ change from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")` because $t^{'}$&lt;$t$ for any $t$.
+We further note here that $K^{'}$&gt;$K$ because annual growth is compressed into a shorter period ($1-NGT$) in `r eqncaps("PaulyMod",display="cite")` when $NGT$&gt;0.
+
+@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at one point, $WP$), replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth), and replacing $K$ with $K^{'}$.
+Furthermore, their modification can be described geometrically (though not algorithmically) in two steps.
+First, the seasonal growth function in `r eqncaps("SomersMod",display="cite")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., the $t^{'}$).
+The growth trajectory is then separated at the $WP$ and a horizontal segment that is $NGT$ units long is inserted.
+This forms a growth trajectory that smoothly transitions into and out of a no-growth period (`r figcaps("PaulyExPlot",display="cite")`).
+
+\
+
+```{r echo=FALSE, results="hide"}
+## Setup
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+WP <- 0.55; NGT <- 0.3
+Linf <- 30; Kpr <- 0.4; t0 <- -0.1
 Pcf <- c(Linf,Kpr,t0,WP,NGT)
-t <- seq(0,5,length.out=299)
+## Get growth trajectory
+t <- seq(-0.1,3.3,length.out=499)
 PL <- paulySC(t,Linf=Pcf)
-plot(PL~t,type="l",lwd=2,ylim=c(0,30),ylab="Length",xlab="Age (years)")
-WPs <- 0:5+WP
+## Make schematic plot
+plot(c(-1,-1),xlim=c(-0.1,3.5),ylim=c(0,22),xaxt="n",
+     ylab="Length",xlab="Age (years)")
+axis(1,0:3)
+## Put on NGT polygons
+# find WPs and Lengths at WPs
+WPs <- 0:2+WP
 LatWPs <- paulySC(WPs,Linf=Pcf)
-segments(WPs,0,WPs,LatWPs,lty=3,lwd=2,col="red")
-segments(WPs+NGT,0,WPs+NGT,LatWPs,lty=3,lwd=2,col="blue")
-legend("topleft",c("WP=0.8","NGT=0.4"),bty="n",inset=-0.02,text.col=c("red","blue"))
+# find start and end of NGT
+NGTs <- WPs-NGT/2
+NGTe <- WPs+NGT/2
+# place polygons
+for (i in 1:length(NGTs)) polygon(c(NGTs[i],NGTs[i],NGTe[i],NGTe[i]),c(0,22,22,0),
+                                  col=col2rgbt("black",1/20),border=NA)
+## Put on the growth trajectory
+lines(PL~t,lwd=2)
+## Place an open point at the WP
+points(WPs,LatWPs,pch=3,col="gray50")
+## Add some important values to the age axis
+axis(1,NGTs,tcl=-0.2)
+axis(1,NGTe,tcl=-0.2)
+## Add the t-prime labels
+axis(1,at=c(0:3,NGTs,NGTe),labels=NA,tcl=0.2)
+axis(1,at=0:3,labels=c(NA,0.7,1.4,2.1),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=NGTs,labels=c(0.4,1.1,1.8),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=NGTe,labels=c(0.4,1.1,1.8),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=3.3,labels="<= t",tick=FALSE)
+axis(1,at=3.3,labels="<= t'",tick=FALSE,line=-1.7,lwd=0)
+
+figcaps("PaulyExPlot",paste0("Example growth trajectory of ",eqncaps('vbMain',display='cite')," using ",eqncaps('PaulyMod',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5,", and $NGT$=",NGT,".
+Each no-growth period is marked by a light gray polygon and each $WP$ is marked by the gray plus symbol.
+The ages adjusted for the $NGT$ (i.e., $t^{'}$) are shown above the x-axis."))
 ```
+
+`r figcaps("PaulyExPlot")`
+
+\
+
+@Paulyetal1992 provided a "diskette" that contained a computer program to estimat the parameters of `eqncaps("PaulyMod",display="cite")`.
+The diskette is difficult to obtain and the source code is no longer available (D.
+Pauly, pers.
+comm.).
+@Paulyetal1992 did describe the operations performed by their program, but their is no description of how $t^{'}$ is operationalized.
+This is an important step in fitting `eqncaps("PaulyMod",display="cite")` as $t^{'}$ is a function of $t$, but also of $NGT$ and $t_{s}$ (or $WP$), which are parameters to be estimated during the model-fitting process.
+In other words, the values for $t^{'}$ will change with each iteration in the non-linear model-fitting process.
+
+The objectives of this note are to (i) describe a function for calculating $t^{'}$; (ii) slightly modify `eqncaps("PaulyMod",display="cite")` to replace $t_{s}$ with $WP$, which has a more useful biological meaning and eases the calculation of $t^{'}$; (iii) provide open-source code for fitting this modified function; and (iv) demonstrate the use of the code for fitting length-at-age data.
+
+# The Modified Model
+
+The $WP$ value is equal to $t_{s}+0.5$ because of the shape of the sin function used to model the oscillations.
+Thus, `r eqncaps("PaulyMod",display="cite")` can be simply modified to include $WP$ rather than $t_{s}$ as
+
+\
+
+$$
+\begin{aligned}
+q & =K^{'}(t^{'}-t_{0}) \\
+  & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-WP-0.5)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`} \\
+  & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-WP-0.5)\right)
+\end{aligned}  
+$$
+
+\
+
+Note that $WP$ in `r eqncaps("PaulyMod2",display="cite")` is the center of the no-growth period (`r figcaps("PaulyEx",display="cite")`).
+Hereafter, `r eqncaps("PaulyMod2",display="cite")` will be used as $WP$ is easier to visualize and to interpret than $t_{s}$.
 
@@
 ```
+
+## References

### 2016-04-14 manuscript_v0.Rmd
again

@@
 ## internal function to compute t-prime
-## 
 ################################################################################
 iCalc_tpr <- function(t,WP,NGT) {
-  ## First NGT stats at WP, use this to shift each age such that
+  ## No-growth period starts at WP-NGT/2.  Use this to shift each age such that
   ##    first NGT would correspond to a value of 0
@@
   ## Find fraction of year on this new time scale for each adjusted age
+  ##    by substracting whole number age (e.g, 2.79-2)
   tmp.t2 <- tmp.t-floor(tmp.t)
-  ## Adjust this fraction for no growth
+  ## Adjust this fraction for no growth ... with the time adjustment no-growth
+  ##    is from 0 to NGT in each "year" ... tprime will not increase in NGT so
+  ##    replace all of these fractions with zeroes.  Above NGT, the tprime must
+  ##    have the NGT removed from it (e.g., if NGT=0.4, then 0.79-0.4 is the
+  ##    fraction of the year for which growth occurred.)
   for (i in 1:length(tmp.t2)) {
@@
   }
+  ## Now get the number of completed growth periods (i.e., floor (tmp.t)) times
+  ##    the length of a growth period (1-NGT) to get the total growth periods
+  ##    completed prior to that year and then add on the fractional piece
   tmp.t <- floor(tmp.t)*(1-NGT)+tmp.t2
@@
 ## Main Function
-##
 ##   Linf, t0 as usual
 ##   Kpr = K-prime as defined in Pauly et al. (units are diff than usual K)
-##   WP = "Winter Period" (point where growth=0) = ts+0.5 (ts from Pauly et al.)
-##   NGT = "No Growth Time" = "fraction of a year where no growth occurs
-##
+##   WP = "Winter Period" (middle point of no-growth period)
+##      = ts+0.5 (ts from Pauly et al.)
+##   NGT = "No Growth Time" = "fraction of a year where no growth occurs"
 ##   tpr = "t-prime" = actual age (t) minus cumulative NGT prior to t
-##   Q is as defined in Pauly et al.
-##   qt and qtr are intermediate values to make Pauly SC look similar to the
-##     Somers and Somers2 models from vbFuns() in FSA
-##
-##   The final line is basically Equation 4 from Pauly et al.
 ################################################################################
@@
   tpr <- iCalc_tpr(t,WP,NGT)
-  Q <- (2*pi)/(1-NGT)
-  qt <- (Kpr/Q)*sin(Q*(tpr-WP-0.5))
-  qt0 <- (Kpr/Q)*sin(Q*(t0-WP-0.5))
-  Linf*(1-exp(-Kpr*(tpr-t0)-qt+qt0))
+  q <- Kpr*(tpr-t0) +
+    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(tpr-WP-0.5)) -
+    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(t0-WP-0.5))
+  Linf*(1-exp(-q))
 }
@@
 
-The mean length-at-age for many animals (e.g., ) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).
+The mean length-at-age for many animals (e.g., CITATIONS) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).
 The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as
@@
 
-Many animals exhibits seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply ().
-`r eqncaps("vbMod",display="cite")` of the traditional VBGF has been modified, usually with a sin function, to model these seasonal oscillations in growth.
+Many animals exhibits seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (CITATIONS).
+`r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sin function, to model these seasonal oscillations in growth.
 The most popular of these modifications is from @HoenigChoudarayHanumara1982 and @Somers1988 (and carefully reiterated in @GarciaBerthouetal2012), and uses
@@
 If $C$=0, then there is no seasonal oscillation and `r eqncaps("SomersMod",display="cite")` reduces to  `r eqncaps("vbMain2",display="cite")` and the typical VBGF (`r figcaps("SomersExPlot",display="cite")`).
-If $C$=1, then growth completely stops once a year at the "winter-point" ($WP=t_{s}+0.5$), whereas values of 0&gt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).
+If $C$=1, then growth completely stops once a year at the "winter-point" ($WP$), whereas values of 0&lt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).
 
@@
 
-figcaps("SomersExPlot",paste0("Example growth trajectory of ",eqncaps('vbMain',display='cite')," using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
+figcaps("SomersExPlot",paste0("Example VBGF using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
 ```
@@
 Values of $C$&gt;1 (or &lt;0) in  `r eqncaps("SomersMod",display="cite")` allow seasonal decreases in mean length-at-age (`r figcaps("SomersExPlot",display="cite")`).
-A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal crease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].
-To address this issue, @Paulyetal1992 modified `r eqncaps("SomersMod")` to include a true seasonal no-growth period with a smooth transition of the modeled mean length-at-age into and out of the no-growth period.
+A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].
+To address this issue, @Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period with a smooth transition of the modeled mean length-at-age into and out of the no-growth period.
 Specifically, their modification is
@@
 
-\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (in fractions of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 introduced $K^{'}$ to `r eqncaps("PaulyMod",display="cite")` because the units of $K$ change from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")` because $t^{'}$&lt;$t$ for any $t$.
-We further note here that $K^{'}$&gt;$K$ because annual growth is compressed into a shorter period ($1-NGT$) in `r eqncaps("PaulyMod",display="cite")` when $NGT$&gt;0.
+\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 introduced $K^{'}$ to `r eqncaps("PaulyMod",display="cite")` because the units of $K$ change from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")` due to $t^{'}$&lt;$t$ for any $t$.
+We further note that $K^{'}$&gt;$K$ because annual growth is compressed into a shorter period ($1-NGT$) in `r eqncaps("PaulyMod",display="cite")` when $NGT$&gt;0.
 
-@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at one point, $WP$), replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth), and replacing $K$ with $K^{'}$.
-Furthermore, their modification can be described geometrically (though not algorithmically) in two steps.
-First, the seasonal growth function in `r eqncaps("SomersMod",display="cite")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., the $t^{'}$).
+@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$), replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth), and replacing $K$ with $K^{'}$.
+Their modification can be described geometrically (though not algorithmically) in two steps.
+First, the seasonal growth function in `r eqncaps("SomersMod",display="cite")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).
 The growth trajectory is then separated at the $WP$ and a horizontal segment that is $NGT$ units long is inserted.
@@
 
-figcaps("PaulyExPlot",paste0("Example growth trajectory of ",eqncaps('vbMain',display='cite')," using ",eqncaps('PaulyMod',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5,", and $NGT$=",NGT,".
+figcaps("PaulyExPlot",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," or ",eqncaps('PaulyMod2',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $NGT$=",NGT,", and $t_{s}$=",WP-0.5," or $WP$=",WP,".
 Each no-growth period is marked by a light gray polygon and each $WP$ is marked by the gray plus symbol.
@@
 
-@Paulyetal1992 provided a "diskette" that contained a computer program to estimat the parameters of `eqncaps("PaulyMod",display="cite")`.
-The diskette is difficult to obtain and the source code is no longer available (D.
+@Paulyetal1992 provided a "diskette" that contained a computer program to estimate the parameters of `r eqncaps("PaulyMod",display="cite")`.
+The diskette is difficult (at best) to obtain and the source code is no longer available (D.
 Pauly, pers.
 comm.).
-@Paulyetal1992 did describe the operations performed by their program, but their is no description of how $t^{'}$ is operationalized.
-This is an important step in fitting `eqncaps("PaulyMod",display="cite")` as $t^{'}$ is a function of $t$, but also of $NGT$ and $t_{s}$ (or $WP$), which are parameters to be estimated during the model-fitting process.
+@Paulyetal1992 did describe the operations performed by their program, but their is no description of how $t^{'}$ was operationalized.
+This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.
 In other words, the values for $t^{'}$ will change with each iteration in the non-linear model-fitting process.
 
-The objectives of this note are to (i) describe a function for calculating $t^{'}$; (ii) slightly modify `eqncaps("PaulyMod",display="cite")` to replace $t_{s}$ with $WP$, which has a more useful biological meaning and eases the calculation of $t^{'}$; (iii) provide open-source code for fitting this modified function; and (iv) demonstrate the use of the code for fitting length-at-age data.
+The objectives of this note are to (i) describe a function for calculating $t^{'}$ that may be used in the model-fitting process; (ii) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while providing a more meaningful parameter; (iii) provide an (open-source) algorithm for the modified function that can be used in model-fitting; and (iv) demonstrate the use of the modified function for fitting length-at-age data.
 
@@
 
-The $WP$ value is equal to $t_{s}+0.5$ because of the shape of the sin function used to model the oscillations.
-Thus, `r eqncaps("PaulyMod",display="cite")` can be simply modified to include $WP$ rather than $t_{s}$ as
+The $WP$ value is equal to $t_{s}+0.5$ because of the shape of the sin function used to model the growth oscillations.
+By simple substitution, `r eqncaps("PaulyMod",display="cite")` is modified to include $WP$ rather than $t_{s}$.
 
@@
 
-Note that $WP$ in `r eqncaps("PaulyMod2",display="cite")` is the center of the no-growth period (`r figcaps("PaulyEx",display="cite")`).
-Hereafter, `r eqncaps("PaulyMod2",display="cite")` will be used as $WP$ is easier to visualize and to interpret than $t_{s}$.
+Note that in `r eqncaps("PaulyMod2",display="cite")` $WP$ is the center of the no-growth period (`r figcaps("PaulyEx",display="cite")`).
 

### 2016-04-15 manuscript_v0.Rmd
Starting to describe simulations in the ms

@@
 ---
-title: "Testing new Pauly Cessational Growth Function"
+title: "Revisiting the von Bertalanffy Seasonal Cessational Growth Function of Pauly et al. (1992)"
 author: "Derek H. Ogle"
@@
 library(FSAsim)
+library(knitr)
+
+source("iCalc_tpr.R")
+source("VBSCGF.R")
 
@@
 
-```
-
-```{r maketprFun, echo=FALSE}
-################################################################################
-## internal function to compute t-prime
-################################################################################
-iCalc_tpr <- function(t,WP,NGT) {
-  ## No-growth period starts at WP-NGT/2.  Use this to shift each age such that
-  ##    first NGT would correspond to a value of 0
-  tshift <- WP-NGT/2
-  tmp.t <- t-tshift
-  ## Find fraction of year on this new time scale for each adjusted age
-  ##    by substracting whole number age (e.g, 2.79-2)
-  tmp.t2 <- tmp.t-floor(tmp.t)
-  ## Adjust this fraction for no growth ... with the time adjustment no-growth
-  ##    is from 0 to NGT in each "year" ... tprime will not increase in NGT so
-  ##    replace all of these fractions with zeroes.  Above NGT, the tprime must
-  ##    have the NGT removed from it (e.g., if NGT=0.4, then 0.79-0.4 is the
-  ##    fraction of the year for which growth occurred.)
-  for (i in 1:length(tmp.t2)) {
-    if (tmp.t2[i]<=NGT) tmp.t2[i] <- 0
-    else tmp.t2[i] <- tmp.t2[i]-NGT
-  }
-  ## Now get the number of completed growth periods (i.e., floor (tmp.t)) times
-  ##    the length of a growth period (1-NGT) to get the total growth periods
-  ##    completed prior to that year and then add on the fractional piece
-  tmp.t <- floor(tmp.t)*(1-NGT)+tmp.t2
-  ## Shift back to get tprime (why is NGT needed here)
-  tpr <- tmp.t+tshift+NGT
-}
-```
-
-```{r makePaulyFun, echo=FALSE}
-################################################################################
-## Main Function
-##   Linf, t0 as usual
-##   Kpr = K-prime as defined in Pauly et al. (units are diff than usual K)
-##   WP = "Winter Period" (middle point of no-growth period)
-##      = ts+0.5 (ts from Pauly et al.)
-##   NGT = "No Growth Time" = "fraction of a year where no growth occurs"
-##   tpr = "t-prime" = actual age (t) minus cumulative NGT prior to t
-################################################################################
-
-paulySC <- function(t,Linf,Kpr=NULL,t0=NULL,WP=NULL,NGT=NULL) {
-  if (length(Linf)==5) { Kpr <- Linf[[2]]; t0 <- Linf[[3]]
-                         WP <- Linf[[4]]; NGT <- Linf[[5]]
-                         Linf <- Linf[[1]] }
-  tpr <- iCalc_tpr(t,WP,NGT)
-  q <- Kpr*(tpr-t0) +
-    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(tpr-WP-0.5)) -
-    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(t0-WP-0.5))
-  Linf*(1-exp(-q))
-}
 ```
@@
 
-\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 introduced $K^{'}$ to `r eqncaps("PaulyMod",display="cite")` because the units of $K$ change from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")` due to $t^{'}$&lt;$t$ for any $t$.
-We further note that $K^{'}$&gt;$K$ because annual growth is compressed into a shorter period ($1-NGT$) in `r eqncaps("PaulyMod",display="cite")` when $NGT$&gt;0.
+\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 noted that the units on $K$ changed from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")`.
+To eliminate confusion he suggested using $K^{'}$ in `r eqncaps("PaulyMod",display="cite")`, as we do here.
 
-@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$), replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth), and replacing $K$ with $K^{'}$.
+@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth and noting that $K^{'}$ only operates in this shorter period).
 Their modification can be described geometrically (though not algorithmically) in two steps.
@@
 WP <- 0.55; NGT <- 0.3
-Linf <- 30; Kpr <- 0.4; t0 <- -0.1
+Linf <- 30; K <- 0.3; Kpr <- K/(1-NGT); t0 <- 0
 Pcf <- c(Linf,Kpr,t0,WP,NGT)
@@
 t <- seq(-0.1,3.3,length.out=499)
-PL <- paulySC(t,Linf=Pcf)
+PL <- VBSCGF(t,Linf=Pcf)
+## Calculate WPs and start and end of NGT
+# find WPs and Lengths at WPs
+WPs <- 0:2+WP
+LatWPs <- VBSCGF(WPs,Linf=Pcf)
+# find start and end of NGT
+NGTs <- WPs-NGT/2
+NGTe <- WPs+NGT/2
 ## Make schematic plot
@@
      ylab="Length",xlab="Age (years)")
-axis(1,0:3)
 ## Put on NGT polygons
-# find WPs and Lengths at WPs
-WPs <- 0:2+WP
-LatWPs <- paulySC(WPs,Linf=Pcf)
-# find start and end of NGT
-NGTs <- WPs-NGT/2
-NGTe <- WPs+NGT/2
-# place polygons
 for (i in 1:length(NGTs)) polygon(c(NGTs[i],NGTs[i],NGTe[i],NGTe[i]),c(0,22,22,0),
@@
 lines(PL~t,lwd=2)
-## Place an open point at the WP
+## Place a point at the WP
 points(WPs,LatWPs,pch=3,col="gray50")
 ## Add some important values to the age axis
+axis(1,0:3)
 axis(1,NGTs,tcl=-0.2)
@@
 ## Add the t-prime labels
+# makes inside ticks
 axis(1,at=c(0:3,NGTs,NGTe),labels=NA,tcl=0.2)
-axis(1,at=0:3,labels=c(NA,0.7,1.4,2.1),tcl=0.2,line=-1.7,lwd=0)
-axis(1,at=NGTs,labels=c(0.4,1.1,1.8),tcl=0.2,line=-1.7,lwd=0)
-axis(1,at=NGTe,labels=c(0.4,1.1,1.8),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=1:3,labels=iCalc_tpr(1:3,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=NGTs,labels=iCalc_tpr(NGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=NGTe,labels=iCalc_tpr(NGTe,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
 axis(1,at=3.3,labels="<= t",tick=FALSE)
@@
 comm.).
-@Paulyetal1992 did describe the operations performed by their program, but their is no description of how $t^{'}$ was operationalized.
+@Paulyetal1992 did describe the operations performed by their program, but there is no description of how $t^{'}$ was operationalized.
 This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.
@@
 
-The objectives of this note are to (i) describe a function for calculating $t^{'}$ that may be used in the model-fitting process; (ii) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while providing a more meaningful parameter; (iii) provide an (open-source) algorithm for the modified function that can be used in model-fitting; and (iv) demonstrate the use of the modified function for fitting length-at-age data.
+Therefore, the objectives of this note are to (i) describe a function for calculating $t^{'}$ that may be used in the model-fitting process; (ii) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while providing a more meaningful parameter; (iii) provide an (open-source) algorithm for the modified function that can be used in model-fitting; and (iv) demonstrate the use of the modified function for fitting length-at-age data.
 
@@
 
-The $WP$ value is equal to $t_{s}+0.5$ because of the shape of the sin function used to model the growth oscillations.
+In `r eqncaps("SomersMod",display="cite")`, $WP=t_{s}+\frac{1}{2}$ because the sine function has a period of one year.
+The growth period is compressed in `r eqncaps("PaulyMod",display="cite")` to be $1-NGT$.
+Thus, the start of the no-growth period ($SNG$) is $t_{s}+\frac{1-NGT}{2}$.
+The center of the no-growth period is then $SNG+\frac{NGT}{2}$ or $t_{s}+\frac{1}{2}$.
+Thus, $WP=t_{s}+\frac{1}{2}$ is the center of the no-growth period for `r eqncaps("PaulyMod",display="cite")`.
+
 By simple substitution, `r eqncaps("PaulyMod",display="cite")` is modified to include $WP$ rather than $t_{s}$.
@@
 q & =K^{'}(t^{'}-t_{0}) \\
-  & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-WP-0.5)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`} \\
-  & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-WP-0.5)\right)
+  & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-WP-\frac{1}{2})\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`} \\
+  & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-WP-\frac{1}{2})\right)
 \end{aligned}  
@@
 
-Note that in `r eqncaps("PaulyMod2",display="cite")` $WP$ is the center of the no-growth period (`r figcaps("PaulyEx",display="cite")`).
+As noted by @Paulyetal1992 the calculation of $t^{'}$ depends on the observed age ($t$) and the cumulative no-growth time prior to $t$.
+In practice, the calculation of $t^{'}$ also depends on the position of the no-growth period within a year.
+Here, the position of the no-growth period is defined by the $SNG$.
+However, because the $SNG$ has a one-to-one relation with $WP$ and $t_{s}$ it can also be defined relative to these parameters.
 
-# Methods
+The following algorithm is used to convert from observed ages ($t$) to ages adjusted for cumulative $NGT$ prior to age $t$ ($t^{'}$).
+
+1.
+Shift the age ($t$) by subtracting the start of the no-growth period ($SNG$) from $t$, such that a whole number age represents the start of a no-growth period.
+For example, if $SNG$=0.4, then $t$=2.4 will become 2.0 and $t$=2.9 will become 2.5.
+1.
+Subtract the whole number age from the shifted age such that the remaining decimal represents the fraction of a shifted year.
+For example, a 0 will result if the shifted age is 2.0 and a 0.5 will result if the shifted age is 2.5.
+1. Substract the $NGT$ from each of the values from the previous step.
+1.
+If the value from the previous step is negative, then the age is within the no-growth period and the negative value should be replaced with a zero.
+Otherwise, the positive value represents the fraction of a growth period completed.
+1. Add the value from the previous step to the total growth time completed (i.e., the product of the number of growth periods completed and the length of the growth period ($1-NGT$)).
+1. Compute $t{'}$ by adding back the $SNG$ that was subtracted in Step 1.
+
+Further examples of $t^{'}$ values relative to $t$ values are shown in `r figcaps("PaulyExPlot",display="cite")`.
+This algorithm for computing $t^{'}$ is implemented in an R [@R2016] function as shown in Appendix 1.
+With this, `r eqncaps("paulyMod2",display="cite")` is easly implemented as an R function as shown in Appendix 2.
+
+# Fitting the Modified Model
 ## Simulated Data
-
-```{r echo=FALSE}
-WP <- 0.8
-NGT <- 0.5
-gper <- c(WP+NGT/2-1,WP-NGT/2)*365
-## Set some growth parameters
-Linf <- 30; K <- 0.3; t0 <- 0
-## Actually generate the data (with little variability)
-d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.01,SD=0.01,paramCV=0.01,
-                growth.per=gper,sample.per=c(1,365),lendigs=1)
+Fish ages (including fractions of a year) and lengths at capture were simulated with vbDataGen() from the FSAsims package (CITATION) in R.
+In this application, this functions simulates seasonal fish growth with the following steps:
+```{r echo=FALSE, results='hide'}
+## Set some common parameters
+Linf <- 30; K <- 0.3; t0 <- -0.1
+## Set bounds for model-fitting
+lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
+uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
 ```
 
-## R Implementation
-
-# Model Fitting
-## Simulated Data
-
-```{r echo=FALSE, message=FALSE}
-lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
-uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
-## Try fitting the PaulySC model
-sv <- list(Linf=32,K=0.32,t0=0,WP=0.8,NGT=0.1)
-fitP1 <- nls(lenCap~paulySC(ageFracY,Linf,K,t0,WP,NGT),data=d1,
-             start=sv,lower=lwrbnd,upper=uprbnd,algorithm="port",control=list(maxiter=100))
-## Make a diagnostic plot
-plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5))
-curve(paulySC(x,coef(fitP1)),from=1,to=6,col="red",lwd=3,add=TRUE)
-cfs <- coef(fitP1)
+```{r echo=FALSE, results="hide", message=FALSE}
+## Setup
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+## Set some specific growth parameters
+# fairly typical north temperate situation
+WP <- 0.8; NGT <- 0.5
+gper <- c(WP+NGT/2-1,WP-NGT/2)*365
+## Actually generate the data (with very little variability)
+d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.001,SD=0.001,paramCV=0.001,
+                growth.per=gper,sample.per=c(1,365),lendigs=1)
+## Try fitting the VBSCGF model
+sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
+fitP1 <- nls(lenCap~VBSCGF(ageFracY,Linf,K,t0,WP,NGT),data=d1,
+             start=sv,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+( cfs <- coef(fitP1) )
+## A plot
+plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/3),
+     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
+curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
 WPs <- 1:5+cfs[["WP"]]
-LatWPs <- paulySC(WPs,Linf=cfs)
-segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col="red")
+LatWPs <- VBSCGF(WPs,Linf=cfs)
+arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
 SNGTs <- WPs-cfs[["NGT"]]/2
 ENGTs <- WPs+cfs[["NGT"]]/2
-segments(SNGTs,LatWPs-1,SNGTs,LatWPs+1,lty=3,lwd=2,col="blue")
-segments(ENGTs,LatWPs-1,ENGTs,LatWPs+1,lty=3,lwd=2,col="blue")
-legend("topleft",paste(c("WP","NGT"),c(WP,NGT),sep="="),
-       bty="n",inset=-0.02,text.col=c("red","blue"))
+arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)
 
-cbind(Est=coef(fitP1),confint(fitP1))
-```
-```{r echo=FALSE}
-plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5),xlim=c(1.7,1.9),ylim=c(13.425,13.43))
-curve(paulySC(x,coef(fitP1)),from=1,to=6,col="red",lwd=3,add=TRUE,n=1001)
-segments(WPs,LatWPs-2,WPs,LatWPs+2,lty=3,lwd=2,col="red")
-segments(SNGTs,LatWPs-1,SNGTs,LatWPs+1,lty=3,lwd=2,col="blue")
-segments(ENGTs,LatWPs-1,ENGTs,LatWPs+1,lty=3,lwd=2,col="blue")
-
+tmp <- c("Linf=","K=","t0=","WP=","NGT=")
+legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP,NGT)),
+       bty="n",inset=-0.02,cex=0.8)
+cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
+legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
 ```
 
-```{r echo=FALSE, message=FALSE}
-## Try fitting the Somers model
-vbSO <- vbFuns("Somers2")
-fitSO1 <- nls(lenCap~vbSO(ageFracY,Linf,K,t0,C,WP),data=d1,
-               start=list(Linf=32,K=0.32,t0=0,C=0.8,WP=0.8),
-               lower=c(Linf=0,K=0,t0=-Inf,C=-1,WP=0),
-               upper=c(Linf=Inf,K=Inf,t0=Inf,C=2,WP=1),
-               algorithm="port",control=list(maxiter=100),trace=FALSE)
-plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/5))
-curve(vbSO(x,coef(fitSO1)),from=1,to=6,col="red",lwd=2,add=TRUE)
-cbind(coef(fitSO1),confint(fitSO1))
+```{r echo=FALSE, results="hide", message=FALSE}
+## Setup
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+## Set some specific growth parameters
+# fairly typical north temperate situation
+WP <- 0.8; NGT <- 0.5
+gper <- c(WP+NGT/2-1,WP-NGT/2)*365
+## Actually generate the data (with more variability)
+d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.2,SD=0.5,paramCV=0.01,
+                growth.per=gper,sample.per=c(1,365),lendigs=1)
+## Try fitting the VBSCGF model
+sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
+fitP1 <- nls(lenCap~VBSCGF(ageFracY,Linf,K,t0,WP,NGT),data=d1,
+             start=sv,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+( cfs <- coef(fitP1) )
+## A plot
+plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/3),
+     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
+curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
+WPs <- 1:5+cfs[["WP"]]
+LatWPs <- VBSCGF(WPs,Linf=cfs)
+arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
+SNGTs <- WPs-cfs[["NGT"]]/2
+ENGTs <- WPs+cfs[["NGT"]]/2
+arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)
+
+tmp <- c("Linf=","K=","t0=","WP=","NGT=")
+legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP,NGT)),
+       bty="n",inset=-0.02,cex=0.8)
+cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
+legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
+```
+
+```{r echo=FALSE, results="hide", message=FALSE}
+## Setup
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+## Set some specific growth parameters
+# fairly typical north temperate situation
+WP <- 0.8; NGT <- 0.5
+gper <- c(WP+NGT/2-1,WP-NGT/2)*365
+## Actually generate the data (with more variability)
+d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.2,SD=0.5,paramCV=0.01,
+                growth.per=gper,sample.per=c(1,365),lendigs=1)
+## Try fitting the VBSCGF model
+sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
+fitP1 <- nls(lenCap~VBSCGF(ageFracY-0.3,Linf,K,t0,WP,NGT),data=d1,
+             start=sv,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+( cfs <- coef(fitP1) )
+## A plot
+plot(lenCap~I(ageFracY-0.3),data=d1,pch=19,col=col2rgbt("black",1/3),
+     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
+curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
+WPs <- 1:5+cfs[["WP"]]
+LatWPs <- VBSCGF(WPs,Linf=cfs)
+arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
+SNGTs <- WPs-cfs[["NGT"]]/2
+ENGTs <- WPs+cfs[["NGT"]]/2
+arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)
+
+tmp <- c("Linf=","K=","t0=","WP=","NGT=")
+legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP-0.3,NGT)),
+       bty="n",inset=-0.02,cex=0.8)
+cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
+legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
 ```
@@
 
-## Appendix
+# Discussion
 
-```{r}
-<<maketprFun>>
+# Appendices
+## Appendix 1
+
+```{r echo=FALSE, comment=""}
+cat(readLines("iCalc_tpr.R"),sep="\n")
 ```
 
-```{r}
-<<makePaulyFun>>
+## Appendix 2
+
+```{r echo=FALSE, comment=""}
+cat(readLines("VBSCGF.R"),sep="\n")
 ```
 
-## References
+# References

### 2016-04-15 VBSCGF.R

@@ -1,0 +1,19 @@
+################################################################################
+## Main Function
+##   Linf, t0 as usual
+##   Kpr = K-prime as defined in Pauly et al. (units are diff than usual K)
+##   WP = "Winter Period" (middle point of no-growth period) = ts+0.5
+##   NGT = "No Growth Time" = "fraction of a year where no growth occurs"
+##   tpr = "t-prime" = actual age (t) minus cumulative NGT prior to t
+################################################################################
+
+VBSCGF <- function(t,Linf,Kpr=NULL,t0=NULL,WP=NULL,NGT=NULL) {
+  if (length(Linf)==5) { Kpr <- Linf[[2]]; t0 <- Linf[[3]]
+  WP <- Linf[[4]]; NGT <- Linf[[5]]
+  Linf <- Linf[[1]] }
+  tpr <- iCalc_tpr(t,WP,NGT)
+  q <- Kpr*(tpr-t0) +
+    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(tpr-WP+1/2)) -
+    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(t0-WP+1/2))
+  Linf*(1-exp(-q))
+}

### 2016-04-15 iCalc_tpr.R

@@ -1,0 +1,15 @@
+################################################################################
+## internal function to compute t-prime
+################################################################################
+iCalc_tpr <- function(t,WP,NGT) {
+  ## Step 1
+  SNG <- WP-NGT/2
+  tmp.t <- t-SNG
+  ## Step 2 (in parentheses) and Step 3
+  tmp.t2 <- tmp.t-floor(tmp.t) - NGT
+  ## Step 4
+  tmp.t2[tmp.t2<0] <- 0
+  ## Step 5 (in parentheses) and Step 6 (also returns value)
+  (floor(tmp.t)*(1-NGT)+tmp.t2) + SNG
+}
+

### 2016-05-09 manuscript_v0_1.Rmd
Started to add Bonito data

---
title: "Revisiting the von Bertalanffy Seasonal Cessational Growth Function of Pauly et al. (1992)"
author: "Derek H. Ogle"
date: "April 29, 2016"
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 4.5
  html_document:
    fig_height: 3.5
    fig_width: 4.5
csl: american-fisheries-society.csl
bibliography: SeasonalGrowth.bib
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34738422)  # for reproducibility
library(FSA)
library(FSAsim)
library(latex2exp)
library(knitr)

source("iCalc_tpr.R")
source("VBSCGF.R")

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
eqncaps("vbMain")
eqncaps("vbMain2")
eqncaps("SomersMod")
eqncaps("PaulyMod")
eqncaps("PaulyMod2")

vbSO <- vbFuns("Somers2")
```

# Introduction

The mean length-at-age for many animals (e.g., CITATIONS) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).  The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as

$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("vbMain",display="num"),")")`} $$

with

$$ q=K(t-t_{0}) \quad \quad \text{`r paste0("(",eqncaps("vbMain2",display="num"),")")`} $$

where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the asymptotic mean length, $K$ is a measure of the exponential rate of approach to the asymptote [@SchnuteFournier1980], and $t_{0}$ is the theoretical time or age (generally negative) at which the mean length would be zero.

Many animals exhibit seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (CITATIONS).  `r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sine function, to model these seasonal oscillations in growth.  The most popular of these modifications is from @HoenigChoudarayHanumara1982 and @Somers1988 (and carefully reiterated in @GarciaBerthouetal2012), and uses

$$
\begin{aligned}
q= & K(t-t_{0}) \\
   & +\frac{CK}{2\pi}sin(2\pi(t-t_{s}))-\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`}
\end{aligned}  
$$

where $C$ modulates the amplitude of the growth oscillations and corresponds to the proportional decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and `r eqncaps("SomersMod",display="cite")` reduces to  `r eqncaps("vbMain2",display="cite")` and the typical VBGF (`r figcaps("SomersExPlot",display="cite")`).  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP$), whereas values of 0&lt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.55; Linf <- 30; K <- 0.3; t0 <- -0.1
C <- c(0,0.5,1,2)
clr <- col2rgbt(rep("black",4),1/(1:4))
t <- seq(0,3,length.out=299)
plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],WP))~t,type="l",lwd=2,col=clr[1],
     ylim=c(0,20),ylab="Mean Length",xlab="Age (years)",xaxt="n")
axis(1,0:3)
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],WP))~t,lwd=2,col=clr[2])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],WP))~t,lwd=2,col=clr[3])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],WP))~t,lwd=2,col=clr[4])

legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)

figcaps("SomersExPlot",paste0("Example VBGF using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
```

`r figcaps("SomersExPlot")`

\

Values of $C$&gt;1 (or &lt;0) in  `r eqncaps("SomersMod",display="cite")` allow seasonal decreases in mean length-at-age (`r figcaps("SomersExPlot",display="cite")`).  A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].  @Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period where mean length was not allowed to decrease and that included a smooth transition of the modeled mean length-at-age into and out of the no-growth period.  Specifically, their modification is

\

$$
\begin{aligned}
q= & K^{'}(t^{'}-t_{0}) \\
   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-t_{s})\right)-\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-t_{s})\right) \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`}
\end{aligned}  
$$

\

\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$" [@Paulyetal1992].  Furthermore,  @Paulyetal1992 noted that the units of $K$ changed from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")`.  To eliminate confusion, they suggested using $K^{'}$ in `r eqncaps("PaulyMod",display="cite")`, as we do here.

@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth and noting that $K^{'}$ only operates during this shorter period).  Their modification may be described geometrically (though not algorithmically) in two steps.  First, the seasonal growth function in `r eqncaps("SomersMod",display="cite")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).  The growth trajectory is then separated at each $WP$ and horizontal segments that are $NGT$ units long are inserted at these points.  This forms a growth trajectory that smoothly transitions into and out of the no-growth periods (`r figcaps("PaulyExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
WP <- 0.55; NGT <- 0.3
Linf <- 30; K <- 0.3; Kpr <- K/(1-NGT); t0 <- 0
Pcf <- c(Linf,Kpr,t0,WP,NGT)
## Get growth trajectory
t <- seq(-0.1,3.3,length.out=499)
PL <- VBSCGF(t,Linf=Pcf)
## Calculate WPs and start and end of NGT
# find WPs and Lengths at WPs
WPs <- 0:2+WP
LatWPs <- VBSCGF(WPs,Linf=Pcf)
# find start and end of NGT
NGTs <- WPs-NGT/2
NGTe <- WPs+NGT/2
## Make schematic plot
plot(c(-1,-1),xlim=c(-0.1,3.5),ylim=c(0,22),xaxt="n",
     ylab="Length",xlab="Age (years)")
## Put on NGT polygons
for (i in 1:length(NGTs)) polygon(c(NGTs[i],NGTs[i],NGTe[i],NGTe[i]),c(0,22,22,0),
                                  col=col2rgbt("black",1/20),border=NA)
## Put on the growth trajectory
lines(PL~t,lwd=2)
## Place a point at the WP
points(WPs,LatWPs,pch=3,col="gray50")
## Add some important values to the age axis
axis(1,0:3)
axis(1,NGTs,tcl=-0.2)
axis(1,NGTe,tcl=-0.2)
## Add the t-prime labels
# makes inside ticks
axis(1,at=c(0:3,NGTs,NGTe),labels=NA,tcl=0.2)
axis(1,at=1:3,labels=iCalc_tpr(1:3,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=NGTs,labels=iCalc_tpr(NGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=NGTe,labels=iCalc_tpr(NGTe,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=3.3,labels="<= t",tick=FALSE)
axis(1,at=3.3,labels="<= t'",tick=FALSE,line=-1.7,lwd=0)

figcaps("PaulyExPlot",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," or ",eqncaps('PaulyMod2',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",formatC(Kpr,format="f",digits=4),", $t_{0}$=",t0,", $NGT$=",NGT,", and $t_{s}$=",WP-0.5," or $WP$=",WP,".  The no-growth periods are marked with gray bars and each $WP$ is marked with a gray plus symbol.  The ages adjusted for the $NGT$ (i.e., $t^{'}$) are shown above the x-axis."))
```

`r figcaps("PaulyExPlot")`

\

@Paulyetal1992 provided a "diskette" that contained a computer program to estimate the parameters of `r eqncaps("PaulyMod",display="cite")`.  The diskette is difficult (at best) to obtain and the source code is no longer available (D. Pauly, pers. comm.).  @Paulyetal1992 did describe the operations performed by their program, but there is no description of how $t^{'}$ was operationalized.  This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.  In other words, the values for $t^{'}$ change with each iteration of the non-linear model-fitting process.

Therefore, the objectives of this note are to (i) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while providing a more meaningful parameter; (ii) describe a function for calculating $t^{'}$ that may be used in the model-fitting process; (iii) provide an (open-source) algorithm for the modified function that can be used in model-fitting; and (iv) demonstrate the use of the modified function for fitting length-at-age data.

# The Modified Seasonal Cessation Growth Function

In `r eqncaps("SomersMod",display="cite")`, $WP=t_{s}+\frac{1}{2}$ because the sine function has a period of one year.  The growth period is compressed in `r eqncaps("PaulyMod",display="cite")` to be $1-NGT$.  Thus, the start of the no-growth period ($SNG$) is $t_{s}+\frac{1-NGT}{2}$.  The center of the no-growth period is then $SNG+\frac{NGT}{2}$ or $t_{s}+\frac{1}{2}$.  Thus, $WP=t_{s}+\frac{1}{2}$ is the center of the no-growth period for `r eqncaps("PaulyMod",display="cite")`.

By simple substitution, `r eqncaps("PaulyMod",display="cite")` may be modified to include $WP$ rather than $t_{s}$.

\

$$
\begin{aligned}
q= & K^{'}(t^{'}-t_{0}) \\
   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t^{'}-WP-\frac{1}{2}\right)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`} \\
   & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t_{0}-WP-\frac{1}{2}\right)\right)
\end{aligned}  
$$

\

As noted by @Paulyetal1992 the calculation of $t^{'}$ depends on the observed age ($t$) and the cumulative no-growth time prior to $t$.  In practice, the calculation of $t^{'}$ also depends on the position of the no-growth period within a year.  Here, the position of the no-growth period is defined by $SNG$ and $NGT$.  However, because the $SNG$ has a one-to-one relation with $WP$ and $t_{s}$, the no-growth period can also be defined relative to these parameters.

The following algorithm is used to convert from observed ages ($t$) to ages adjusted for cumulative $NGT$ prior to age $t$ ($t^{'}$).

1. Shift the age ($t$) by subtracting the start of the no-growth period ($SNG$) from $t$, such that a whole number age represents the start of a no-growth period.  For example, if $SNG$=0.4, then $t$=2.4 will become 2.0 and $t$=2.9 will become 2.5.
1. Subtract the whole number age from the shifted age such that the remaining decimal represents the fraction of a shifted year.  For example, a 0 will result if the shifted age is 2.0 and a 0.5 will result if the shifted age is 2.5.
1. Substract the $NGT$ from the value from the previous step.
1. If the value from the previous step is negative, then the age is within the no-growth period and the negative value should be replaced with a zero.  Otherwise, the positive value represents the fraction of the growth period completed.
1. Add the value from the previous step to the total growth time completed (i.e., the product of the number of growth periods completed and the length of the growth period ($1-NGT$)).
1. Compute $t{'}$ by adding back the $SNG$ that was subtracted in Step 1.

Further examples of $t^{'}$ values relative to $t$ values are shown in `r figcaps("PaulyExPlot",display="cite")`.  This algorithm for computing $t^{'}$ is implemented in an R [@R2016] function as shown in Appendix 1.  With this, `r eqncaps("paulyMod2",display="cite")` is easly implemented as an R function as shown in Appendix 2.

# Fitting the Modified Model
Simple simulated data were initially used to demonstrate the fitting of `r eqncaps("paulyMod2",display="cite")`.  The data were simulated by randomly selecting 200 real numbers from a uniform distribution between 0 and 5 to serve as observed ages, plugging these ages into `r eqncaps("paulyMod2",display="cite")` to compute a mean length for each age, and then adding a random deviate from a normal distribution with a mean of 0 and a standard deviation of $\sigma$ to each mean length to simulate observed individual lengths.  The `nls()` function from R was then used to estimate parameter values for `r eqncaps("paulyMod2",display="cite")` from the simulated data.  The "port" algorithm in `nls()` was used as $L_{\infty}$ and $K^{'}$ were constrained to be positive and $WP$ and $NGT$ were constrained to be between 0 and 1.  One would expect the parameter estimates to be very close to the values used to create the simulated data, as `r eqncaps("paulyMod2",display="cite")` was used to both create and fit the data.

One example of a model fit to data that were simulated to have a "long" NGT, "late" WP, and low individual variability is shown in `r figcaps("PaulyExData1",display="cite")`.  In this example, the parameter estimates are quite close to the parameter values used to simulate the data (the set parameter values are within the confidence intervals for each parameter).

```{r echo=FALSE, results='hide'}
## Set some common parameters
Linf <- 30; K <- 0.3; t0 <- -0.1
## Set bounds for model-fitting
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
```
```{r echo=FALSE, results="hide", message=FALSE}
## Set some specific growth parameters
# fairly typical north temperate situation
WP <- 0.8; NGT <- 0.5
sigma <- 0.3
ages <- runif(200,min=0,max=5)
lens <- VBSCGF(ages,Linf,K,t0,WP,NGT)+rnorm(200,0,sigma)
dB1 <- data.frame(ages,lens)
## Try fitting the VBSCGF model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
fitB1 <- nls(lens~VBSCGF(ages,Linf,K,t0,WP,NGT),data=dB1,
             start=sv,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
cfs <- cbind(Est=coef(fitB1),confint(fitB1))
## A plot
# Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
plot(lens~ages,data=dB1,pch=19,col=col2rgbt("black",1/3),
     xlab="Age (years)",ylab="Length",xlim=c(0,5),ylim=c(0,18))
curve(VBSCGF(x,cfs[,"Est"]),from=0,to=6,lwd=2,add=TRUE)
WPs <- 0:5+cfs["WP","Est"]
LatWPs <- VBSCGF(WPs,Linf=cfs[,"Est"])
arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
SNGTs <- WPs-cfs["NGT","Est"]/2
ENGTs <- WPs+cfs["NGT","Est"]/2
arrows(SNGTs,LatWPs+1.2,ENGTs,LatWPs+1.2,lwd=2,length=0.025,angle=90,code=3)

legend("topleft",c(TeX("\\mathbf{Set Values}"),
                   TeX(paste("$L_{\\infty}$",Linf,sep=" = ")),
                   paste("K'",K,sep=" = "),
                   TeX(paste("$t_{0}",t0,sep=" = ")),
                   paste("WP",WP,sep=" = "),
                   paste("NGT",NGT,sep=" = ")),
        bty="n",inset=-0.02,cex=0.8)
legend("bottomright",c(TeX("\\mathbf{Estimated Values}"),
                       TeX(paste0("$L_{\\infty}$"," = ",
                                  formatC(cfs["Linf","Est"],format="f",digits=1),
                                  " (",formatC(cfs["Linf","2.5%"],format="f",digits=1),
                                  ", ",formatC(cfs["Linf","97.5%"],format="f",digits=1),")")),
                   paste0("K' = ",formatC(cfs["K","Est"],format="f",digits=2),
                                  " (",formatC(cfs["K","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["K","97.5%"],format="f",digits=2),")"),
                   TeX(paste0("$t_{0}$"," = ",
                                  formatC(cfs["t0","Est"],format="f",digits=2),
                                  " (",formatC(cfs["t0","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["t0","97.5%"],format="f",digits=2),")")),
                   paste0("WP = ",formatC(cfs["WP","Est"],format="f",digits=2),
                                  " (",formatC(cfs["WP","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["WP","97.5%"],format="f",digits=2),")"),
                   paste0("NGT = ",formatC(cfs["NGT","Est"],format="f",digits=2),
                                  " (",formatC(cfs["NGT","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["NGT","97.5%"],format="f",digits=2),")")),
        bty="n",cex=0.8)

figcaps("PaulyExData1",paste0("Example VBGF using ",eqncaps('PaulyMod2',display='cite')," with parameters shown in the upper-left corner and $\\sigma$=",sigma,".  Parameter estimates (and 95% confidence intervals) from non-linear regression are shown the lower-right corner.  Each WP is shown by the arrow and each NGT is shown by the horizontal interval bar centered on the WP arrow."))
```

`r figcaps("PaulyExData1")`

\

`r eqncaps("paulyMod2",display="cite")` was fit to simulated data sets that consisted of all combinations of "early" ($WP=0.2$) and "late" ($WP=0.8$) winter points, "short" ($NGT=0.2$) and "long" ($NGT=0.5$) no-growth periods, and "low" ($\sigma=0.3$) and "high" ($\sigma=0.5$) individual variabilities.  These limited simulations suggest that all parameters are consistently well estimated, with the possible exception of $t_{0}$.

\

```{r echo=FALSE, message=FALSE}
WPs <- rep(c(0.8,0.2),each=4)
NGTs <- rep(rep(c(0.5,0.2),each=2),2)
sigmas <- rep(c(0.3,0.5),4)
ages <- runif(200,min=0,max=5)
sv <- list(Linf=32,K=0.32,t0=0,WP=0.5,NGT=0.5)

## Start with results from above
res <- matrix(nrow=8,ncol=18)
res[1,] <- c(WPs[1],NGTs[1],sigmas[1],round(cfs["WP",],2),round(cfs["NGT",],2),round(cfs["Linf",],1),round(cfs["K",],2),round(cfs["t0",],2))
for (i in 2:length(WPs)) {
  tmp <- data.frame(ages,lens=VBSCGF(ages,Linf,K,t0,WPs[i],NGTs[i])+rnorm(200,0,sigmas[i]))
  tmpfit <- nls(lens~VBSCGF(ages,Linf,K,t0,WP,NGT),data=tmp,
                start=sv,lower=lwrbnd,upper=uprbnd,
                algorithm="port",control=list(maxiter=100))
  tmpcfs <- cbind(Est=coef(tmpfit),confint(tmpfit))
  res[i,] <- c(WPs[i],NGTs[i],sigmas[i],round(tmpcfs["WP",],2),round(tmpcfs["NGT",],2),round(tmpcfs["Linf",],1),round(tmpcfs["K",],2),round(tmpcfs["t0",],2))
}
colnames(res) <- c("WPset","NGTset","sigmaset","WPest","WPlci","WPuci","NGTest","NGTlci","NGTuci","Linfest","Linflci","Linfuci","Kest","Klci","Kuci","t0est","t0lci","t0uci")
res
```

## Real Data

@Stewartetal2013 examined the growth of 215 Bonito (*Sarda australis*) sampled from commercial landings.  Detailed methods are described in @Stewartetal2013, but note here that fork lengths (mm) were measured for each fish and ages were the decimal age calculated as the number of opaque zones observed on otolith thin sections plus the proportion of the year after the designated birthdate.  @Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to be less than 1.  Their model fit resulted in the boundary condition of $C=1$, which suggested that Bonito ceased to grow at *at least* one point.  As $C$ is at least equal to 1, this result also suggests that `r eqncaps("paulyMod2",display="cite")` should be fit to these data to determine the length of the no-growth period [@Paulyetal1992].  Thus, `r eqncaps("paulyMod2",display="cite")` is fit to these data here as an example of fitting  `r eqncaps("paulyMod2",display="cite")` to real data.

```{r echo=FALSE, results='hide'}
## Bonito data
bon <- read.csv("Bonito.csv")

################################################################################
## Purely Demonstration
## See if results match that from Stewart et al. (2013)
vbSO <- vbFuns("Somers")
lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=1,ts=1)
svBon <- list(Linf=72,K=0.27,t0=0,C=0.6,ts=0.2)
fitBon1 <- nls(FL~vbSO(Age,Linf,K,t0,C,ts),data=bon,
               start=svBon,lower=lwrbnd,upper=uprbnd,
               algorithm="port",control=list(maxiter=100))
## Linf=71.9, K=0.27, t0=-1.92, C=1, and to=0.09 ... they all match (within rounding)
coef(fitBon1)
################################################################################
```

```{r echo=FALSE, results='hide', message=FALSE}
## Fit new model to Bonito data
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
svBon <- list(Linf=72,K=0.27,t0=0,WP=0.5,NGT=0.2)
fitBon2 <- nls(FL~VBSCGF(Age,Linf,K,t0,WP,NGT),data=bon,
             start=svBon,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
cfBon2 <- cbind(Est=coef(fitBon2),confint(fitBon2))
```

```{r echo=FALSE, results="hide"}
## A plot
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
plot(FL~Age,data=bon,pch=19,col="black",
     xlab="Age (years)",ylab="Length",xlim=c(0,4),ylim=c(0,70))
curve(VBSCGF(x,coef(fitBon2)),from=0,to=4,lwd=4,add=TRUE)
curve(vbSO(x,coef(fitBon1)),from=0,to=4,lwd=2,add=TRUE,col="gray50")
legend("bottomright",c(TeX("\\mathbf{Estimated Values}"),
                       TeX(paste0("$L_{\\infty}$"," = ",
                                  formatC(cfBon2["Linf","Est"],format="f",digits=1),
                                  " (",formatC(cfBon2["Linf","2.5%"],format="f",digits=1),
                                  ", ",formatC(cfBon2["Linf","97.5%"],format="f",digits=1),")")),
                   paste0("K' = ",formatC(cfBon2["K","Est"],format="f",digits=2),
                                  " (",formatC(cfBon2["K","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfBon2["K","97.5%"],format="f",digits=2),")"),
                   TeX(paste0("$t_{0}$"," = ",
                                  formatC(cfBon2["t0","Est"],format="f",digits=2),
                                  " (",formatC(cfBon2["t0","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfBon2["t0","97.5%"],format="f",digits=2),")")),
                   paste0("WP = ",formatC(cfBon2["WP","Est"],format="f",digits=2),
                                  " (",formatC(cfBon2["WP","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfBon2["WP","97.5%"],format="f",digits=2),")"),
                   paste0("NGT = ",formatC(cfBon2["NGT","Est"],format="f",digits=2),
                                  " (",formatC(cfBon2["NGT","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfBon2["NGT","97.5%"],format="f",digits=2),")")),
        bty="n",cex=0.8)

figcaps("PaulyBonito",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod2',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite')," (gray line) superimposed.  The parameter estimates (and 95% confidence intervals) from non-linear regression for ",eqncaps('PaulyMod2',display='cite')," are shown the lower-right corner."))
```

`r figcaps("PaulyBonito")`

\

# Discussion

* General -- fits perfectly simulated data well.
* Assumptions -- WP same time each year and age, NGT same length each year and age
* Model-Fitting -- bounding parameters, more difficult to fit five parameters
* Practical Difference -- not large unless C>>1 and NGT>>0

# Appendices
## Appendix 1

```{r echo=FALSE, comment=""}
cat(readLines("iCalc_tpr.R"),sep="\n")
```

## Appendix 2

```{r echo=FALSE, comment=""}
cat(readLines("VBSCGF.R"),sep="\n")
```

# References

### 2016-05-11 manuscript_v0_2.Rmd
Removed WP and put ts in

---
title: "Revisiting the von Bertalanffy Seasonal Cessational Growth Function of Pauly et al. (1992)"
author: "Derek H. Ogle"
date: "April 29, 2016"
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 4.5
  html_document:
    fig_height: 3.5
    fig_width: 4.5
csl: american-fisheries-society.csl
bibliography: SeasonalGrowth.bib
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34738422)  # for reproducibility
library(FSA)
library(FSAsim)
library(knitr)

source("iCalc_tpr.R")
source("VBSCGF.R")

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
eqncaps("vbMain")
eqncaps("vbMain2")
eqncaps("SomersMod")
eqncaps("PaulyMod")
eqncaps("PaulyMod2")

vbSO <- vbFuns("Somers")
```

# Introduction

The mean length-at-age for many animals (e.g., CITATIONS) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).  The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as

$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("vbMain",display="num"),")")`} $$

with

$$ q=K(t-t_{0}) \quad \quad \text{`r paste0("(",eqncaps("vbMain2",display="num"),")")`} $$

where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the asymptotic mean length, $K$ is a measure of the exponential rate of approach to the asymptote [@SchnuteFournier1980], and $t_{0}$ is the theoretical time or age (generally negative) at which the mean length would be zero.

Many animals exhibit seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (CITATIONS).  `r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sine function, to model these seasonal oscillations in growth.  The most popular of these modifications, from @HoenigChoudarayHanumara1982 and @Somers1988, is

$$
\begin{aligned}
q= & K(t-t_{0}) \\
   & +\frac{CK}{2\pi}sin(2\pi(t-t_{s}))-\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`}
\end{aligned}  
$$

where $C$ modulates the amplitude of the growth oscillations and corresponds to the proportional decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and `r eqncaps("SomersMod",display="cite")` reduces to  `r eqncaps("vbMain2",display="cite")` and the typical VBGF (`r figcaps("SomersExPlot",display="cite")`).  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP$), whereas values of 0&lt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).  Note that $WP=t_{s}+\frac{1}{2}$ because the sine function in `r eqncaps("SomersMod",display="cite")` has a period (i.e., the growth period) of one year.  Some confusion has surrounded the use of `r eqncaps("SomersMod",display="cite")`, though @GarciaBerthouetal2012 carefully clarified its form.

\

```{r echo=FALSE, results="hide"}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
ts <- 0.05; Linf <- 30; K <- 0.3; t0 <- -0.1
C <- c(0,0.5,1,2)
clr <- col2rgbt(rep("black",4),1/(1:4))
t <- seq(0,3,length.out=299)
plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],ts))~t,type="l",lwd=2,col=clr[1],
     ylim=c(0,20),ylab="Mean Length",xlab="Age (years)",xaxt="n")
axis(1,0:3)
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],ts))~t,lwd=2,col=clr[2])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],ts))~t,lwd=2,col=clr[3])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],ts))~t,lwd=2,col=clr[4])

legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)

figcaps("SomersExPlot",paste0("Example VBGF using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",ts," (with $WP$=",ts+0.5,") and four different values of $C$."))
```

`r figcaps("SomersExPlot")`

\

Values of $C$&gt;1 (or &lt;0) in  `r eqncaps("SomersMod",display="cite")` allow seasonal decreases in mean length-at-age (`r figcaps("SomersExPlot",display="cite")`).  A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].  @Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period where mean length was not allowed to decrease and that included a smooth transition of the modeled mean length-at-age into and out of the no-growth period.  Specifically, their modification is

\

$$
\begin{aligned}
q= & K^{'}(t^{'}-t_{0}) \\
   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-t_{s})\right)-\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-t_{s})\right) \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`}
\end{aligned}  
$$

\

\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$" [@Paulyetal1992].  Furthermore,  @Paulyetal1992 noted that the units of $K$ changed from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")`.  To eliminate confusion, they suggested using $K^{'}$ in `r eqncaps("PaulyMod",display="cite")`, as we do here.

@Paulyetal1992 devised  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the seasonal oscillation to the growth period and noting that $K^{'}$ only operates during the growth period).  Their modification may be described geometrically (though not algorithmically) in two steps.  First, the seasonal growth function with $C$=1 in `r eqncaps("SomersMod",display="cite")` is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).  The growth trajectory is then separated at each $WP$ and horizontal segments that are $NGT$ units long are inserted at these points.  This forms a growth trajectory that smoothly transitions into and out of the no-growth periods (`r figcaps("PaulyExPlot",display="cite")`).

\

```{r echo=FALSE, results="hide"}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
## most parametrs from above
Kpr <- 0.35; NGT <- 0.3
Pcf <- c(Linf,Kpr,t0,ts,NGT)
## Get growth trajectory
t <- seq(-0.1,3.3,length.out=499)
PL <- VBSCGF(t,Linf=Pcf)
## Make schematic plot
plot(c(-1,-1),xlim=c(-0.1,3.5),ylim=c(0,22),xaxt="n",
     ylab="Length",xlab="Age (years)")
## Put on NGT polygons
WPs <- 0:2+ts+0.5
LatWPs <- VBSCGF(WPs,Linf=Pcf)
SNGTs <- WPs-NGT/2
ENGTs <- WPs+NGT/2
for (i in 1:length(SNGTs)) polygon(c(SNGTs[i],SNGTs[i],ENGTs[i],ENGTs[i]),c(0,LatWPs[i],LatWPs[i],0),
                                   col=col2rgbt("black",1/20),border=NA)
## Mark the WPs
arrows(WPs,LatWPs-2,WPs,LatWPs-0.2,lwd=2,length=0.1)
## Mark the NGTs
arrows(SNGTs,LatWPs-1.2,ENGTs,LatWPs-1.2,lwd=2,length=0.025,angle=90,code=3)
## Put on the growth trajectory
lines(PL~t,lwd=2)
## Put on the ts values
tss <- 0:3+ts
Lattss <- VBSCGF(tss,Linf=Pcf)
points(tss,Lattss,pch=16,col="gray50",cex=1.1)
## Add some important values to the age axis
axis(1,0:3)
axis(1,SNGTs,tcl=-0.2)
axis(1,ENGTs,tcl=-0.2)
## Add the t-prime labels
# makes inside ticks
axis(1,at=c(0:3,SNGTs,ENGTs),labels=NA,tcl=0.2)
axis(1,at=1:3,labels=iCalc_tpr(1:3,ts,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=SNGTs,labels=iCalc_tpr(SNGTs,ts,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=ENGTs,labels=iCalc_tpr(ENGTs,ts,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=3.3,labels="<= t",tick=FALSE)
axis(1,at=3.3,labels="<= t'",tick=FALSE,line=-1.7,lwd=0)

figcaps("PaulyExPlot",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $NGT$=",NGT,", and $t_{s}$=",ts," (with $WP$=",ts+0.5,").  Each $t_{s}$ is shown by a gray point, $WP$ by a vertical arrow, and no-growth period by the horizontal interval centered on the $WP$ arrow and the gray region that extends to the x-axis.  The ages adjusted for the $NGT$ (i.e., $t^{'}$) are shown above the x-axis."))
```

`r figcaps("PaulyExPlot")`

\

@Paulyetal1992 provided a "diskette" that contained a computer program to estimate the parameters of `r eqncaps("PaulyMod",display="cite")`.  The diskette is difficult (at best) to obtain and the source code is no longer available (D. Pauly, pers. comm.).  @Paulyetal1992 did describe the operations performed by their program, but there is no description of how $t^{'}$ was operationalized.  This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.  Thus, the values for $t^{'}$ change with each iteration of the non-linear model-fitting process.

Therefore, the objectives of this note are to (i) operationalize the calculation of $t^{'}$, (ii) provide an (open-source) algorithm for the calculation of $t^{'}$ and `r eqncaps("PaulyMod",display="cite")` for use in model fitting, and (iii) demonstrate the use of this algorithm.

# Calculating  $t^{'}$

As noted by @Paulyetal1992 the calculation of $t^{'}$ depends on the observed age ($t$) and the cumulative no-growth time prior to $t$.  In practice, the calculation of $t^{'}$ also depends on the position of the no-growth period within a year.  Here, the position of the no-growth period will be defined relative to $WP$ and $NGT$.  The following algorithm may be used to convert from observed ages ($t$) to ages adjusted for cumulative $NGT$ prior to age $t$ ($t^{'}$).

1. Shift the age ($t$) by subtracting the start of the no-growth ($SNG$) period (i.e., $SNG=WP-\frac{NGT}{2}$) from $t$, such that a whole number will represent the start of a no-growth period.  For example, if $SNG$=0.4, then $t$=2.4 will become 2.0 and $t$=2.9 will become 2.5.
1. Subtract the whole number age from the shifted age from Step 1 such that the remaining decimal represents the fraction of a shifted year.  For example, a 0 will result if the shifted age is 2.0 and a 0.5 will result if the shifted age is 2.5.
1. Substract the $NGT$ from the value from the previous step.
1. If the value from the previous step is negative, then the age is within the no-growth period and the negative value should be replaced with a zero.  Otherwise, the positive value represents the amount of time into a growth period.
1. Add the value from the previous step to the total growth time completed (i.e., the product of the number of growth periods completed and the length of the growth period ($1-NGT$)).
1. Compute $t{'}$ by adding back the $SNG$ that was subtracted in Step 1.

Further examples of $t^{'}$ values relative to $t$ values are shown in `r figcaps("PaulyExPlot",display="cite")`.  This algorithm for computing $t^{'}$ is implemented in an R [@R2016] function as shown in Appendix 1.  With this, `r eqncaps("PaulyMod",display="cite")` is easly implemented as an R function as shown in Appendix 2.

# Fitting the Function
## Simulated Data
Simple simulated data are used initially to demonstrate the fitting of `r eqncaps("PaulyMod",display="cite")`.  The data were simulated by randomly selecting 200 real numbers from a uniform distribution between 0 and 5 to serve as observed ages, plugging these ages into `r eqncaps("PaulyMod",display="cite")` to compute a mean length for each age, and then adding a random deviate from a normal distribution with a mean of 0 and a standard deviation of $\sigma$ to each mean length to simulate observed individual lengths.  Eight data sets were simulated for all combinations of "early" ($t_{s}=0.2$) and "late" ($t_{s}=0.8$) maximum growth points, "short" ($NGT=0.2$) and "long" ($NGT=0.5$) no-growth periods, and "low" ($\sigma=0.3$) and "high" ($\sigma=0.5$) individual variabilities.

The `nls()` function from R was then used to estimate parameter values for the nonlinear `r eqncaps("PaulyMod",display="cite")` fit to each simulated data set.  The "port" algorithm was used so that $L_{\infty}$ and $K^{'}$ could be constrained to be positive and $t_{s}$ and $NGT$ could be constrained to be between 0 and 1.  The estimated parameters for each data were consistently very close to the values used to create the data set, with the possible exception of $t_{0}$.  One example, for a "long" $NGT$, "late" $t_{s}$, and low individual variability, is shown in `r figcaps("PaulySimRes",display="cite")`.  The results from fitting `r eqncaps("PaulyMod",display="cite")` to these data sets are not surprising as one would expect the parameter estimates to be very close to the values used to create them, given that `r eqncaps("PaulyMod2",display="cite")` was used to both create and fit the data.

```{r echo=FALSE, results="hide", message=FALSE}
## Used Linf, Kpr, and t0 from above
# fairly typical north temperate situation
tss <- rep(c(0.1,0.5),each=4)
NGTs <- rep(rep(c(0.2,0.4),each=2),2)
sigmas <- rep(c(0.3,0.5),4)

ages <- runif(200,min=0.2,max=5)
lens <- VBSCGF(ages,Linf,Kpr,t0,tss[1],NGT[1])+rnorm(200,0,sigmas[1])
dB1 <- data.frame(ages,lens)
## Try fitting the VBSCGF model
sv <- list(Linf=32,Kpr=0.32,t0=0,ts=0.2,NGT=0.3)
lwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
uprbnd <- c(Linf=Inf,Kpr=Inf,t0=Inf,ts=1,NGT=1)
fitB1 <- nls(lens~VBSCGF(ages,Linf,Kpr,t0,ts,NGT),data=dB1,
             start=sv,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
cfs <- cbind(Est=coef(fitB1),confint(fitB1))
## A plot
# Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
plot(lens~ages,data=dB1,pch=19,col=col2rgbt("black",1/3),
     xlab="Age (years)",ylab="Length",xlim=c(0,5),ylim=c(0,22))
curve(VBSCGF(x,cfs[,"Est"]),from=0,to=6,lwd=2,add=TRUE)
WPs <- 0:5+cfs["ts","Est"]+0.5
LatWPs <- VBSCGF(WPs,Linf=cfs[,"Est"])
arrows(WPs,LatWPs-2,WPs,LatWPs-0.2,lwd=2,length=0.1)
SNGTs <- WPs-cfs["NGT","Est"]/2
ENGTs <- WPs+cfs["NGT","Est"]/2
arrows(SNGTs,LatWPs-1.2,ENGTs,LatWPs-1.2,lwd=2,length=0.025,angle=90,code=3)

figcaps("PaulySimRes",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",Kpr,", $t_{0}$=",t0,", $NGT$=",NGTs[1],", $t_{s}$=",tss[1]," (with $WP$=",ts+0.5,") and $\\sigma$=",sigmas[1],".  Parameter estimates (and 95% confidence intervals) are shown in the first row of ",tblcaps("PaulySimRes",display="cite"),".  Each $WP$ is shown by the vertical arrow and each no-growth period is shown by the horizontal interval bar centered on the $WP$ arrow."))
```

`r figcaps("PaulySimRes")`

\

```{r echo=FALSE, results="hide", message=FALSE}
## Start with results from above
res <- matrix(nrow=8,ncol=18)
res[1,] <- c(tss[1],NGTs[1],sigmas[1],round(cfs["ts",],2),round(cfs["NGT",],2),round(cfs["Linf",],1),round(cfs["Kpr",],2),round(cfs["t0",],2))
for (i in 2:length(tss)) {
  tmp <- data.frame(ages,lens=VBSCGF(ages,Linf,Kpr,t0,tss[i],NGTs[i])+rnorm(200,0,sigmas[i]))
  tmpfit <- nls(lens~VBSCGF(ages,Linf,Kpr,t0,ts,NGT),data=tmp,
                start=sv,lower=lwrbnd,upper=uprbnd,
                algorithm="port",control=list(maxiter=100))
  tmpcfs <- cbind(Est=coef(tmpfit),confint(tmpfit))
  res[i,] <- c(tss[i],NGTs[i],sigmas[i],round(tmpcfs["ts",],2),round(tmpcfs["NGT",],2),round(tmpcfs["Linf",],1),round(tmpcfs["Kpr",],2),round(tmpcfs["t0",],2))
}
colnames(res) <- c("tsset","NGTset","sigmaset","tsest","tslci","tsuci","NGTest","NGTlci","NGTuci","Linfest","Linflci","Linfuci","Kprest","Kprlci","Kpruci","t0est","t0lci","t0uci")

tblcaps("PaulySimRes",paste0("Parameter estimates from fitting ",eqncaps('PaulyMod2',display='cite')," to simulated data sets using $L_{\\infty}$=",Linf,", $K'$=",Kpr,", $t_{0}$=",t0,", and varying values of $t_{s}$, $NGT$, and $\\sigma$."))
```

`r tblcaps("PaulySimRes")`

```{r echo=FALSE, comment=""}
res
```

\

## Real Data

@Stewartetal2013 examined the growth of 215 Bonito (*Sarda australis*) sampled from commercial landings.  Fork lengths (mm) were measured for each fish and decimal ages were recorded as the number of opaque zones observed on otolith thin sections plus the proportion of the year after the designated birthdate (see @Stewartetal2013 for more detailed methods).  @Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to not exceed 1.  Their model fit resulted in the boundary condition of $C=1$, which suggested that Bonito ceased to grow at *at least* one point.  This result suggests that `r eqncaps("PaulyMod2",display="cite")` should be fit to these data to determine the length of the no-growth period [@Paulyetal1992].

```{r echo=FALSE, results='hide', message=FALSE}
## Bonito data
bon <- read.csv("Bonito.csv")

################################################################################
## Purely Demonstration
## See if results match that from Stewart et al. (2013)
lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=1,ts=1)
svBon <- list(Linf=72,K=0.27,t0=0,C=0.6,ts=0.2)
fitBon1 <- nls(FL~vbSO(Age,Linf,K,t0,C,ts),data=bon,
               start=svBon,lower=lwrbnd,upper=uprbnd,
               algorithm="port",control=list(maxiter=100))
## Linf=71.9, K=0.27, t0=-1.92, C=1, and ts=0.09 ... they all match (within rounding)
cfBon1 <- cbind(Est=coef(fitBon1),confint(fitBon1))
################################################################################
```

```{r echo=FALSE, results='hide', message=FALSE}
## Fit new model to Bonito data
lwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
uprbnd <- c(Linf=Inf,Kpr=Inf,t0=Inf,ts=1,NGT=1)
svBon <- list(Linf=72,Kpr=0.27,t0=0,ts=0.5,NGT=0.2)
fitBon2 <- nls(FL~VBSCGF(Age,Linf,Kpr,t0,ts,NGT),data=bon,
             start=svBon,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
cfBon2 <- cbind(Est=coef(fitBon2),confint(fitBon2))
```

`r eqncaps("PaulyMod",display="cite")` fit the Bonito data slightly better than `r eqncaps("SomersMod",display="cite")` with slightly lower residual sums-of-squares (RSS) and Akaike Information Criterion (AIC) values.  The length of the no-growth period was estimated to be `r formatC(cfBon2["NGT","Est"],format="f",digits=3)` or `r formatC(cfBon2["NGT","Est"]*100,format="f",digits=1)`% of the year.  The $t_{s}$ parameters were equal, the $L_\infty$ parameters were similar, but the $t_{0}$ parameters differed somewhat between the two models.  Finally, $K$ from `r eqncaps("SomersMod",display="cite")` was equal to $K'$ from `r eqncaps("PaulyMod2",display="cite")` multiplied by $1-NGT$.

```{r echo=FALSE, results="hide"}
## A plot
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
plot(FL~Age,data=bon,pch=19,col="black",
     xlab="Age (years)",ylab="Fork Length (mm)",xlim=c(0,4),ylim=c(0,70))
curve(VBSCGF(x,coef(fitBon2)),from=0,to=4,lwd=4,add=TRUE)
curve(vbSO(x,coef(fitBon1)),from=0,to=4,lwd=2,add=TRUE,col="gray50")

figcaps("BonitoRes",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod2',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite')," (gray line) superimposed.  The parameter estimates (and 95% confidence intervals) from the model fits are shown in ",tblcaps("BonitoRes",display='cite'),"."))
```

`r figcaps("BonitoRes")`

\

```{r echo=FALSE, results="hide"}
cfBon1 <- rbind(cfBon1,c(sum(summary(fitBon1)$residuals^2),NA,NA),c(AIC(fitBon1),NA,NA))
cfBon2 <- rbind(cfBon2,c(sum(summary(fitBon2)$residuals^2),NA,NA),c(AIC(fitBon2),NA,NA))
rownames(cfBon1)[6:7] <- rownames(cfBon2)[6:7] <- c("RSS","AIC")

tblcaps("BonitoRes",paste0("Parameter estimates, residual sums-of-squares (RSS), and Akaike Information Criterion (AIC) from the fits of ",eqncaps('SomersMod',display='cite')," and ",eqncaps('PaulyMod2',display='cite')," to the Bonito data."))
```

`r tblcaps("BonitoRes")`

```{r echo=FALSE, comment=""}
cat("Somers Model (Equation 3) fit")
print(round(cfBon1,2),na.print="-")
cat("Modified Pauly Model (Equation 6) fit")
print(round(cfBon2,2),na.print="-")
```

\

# Discussion

* General
    * Fits perfectly simulated data well.
    * Parameters from real data seem reasonable
    * Other parameters by maths -- WP=ts+0.5, SNG=WP-NGT/2, K=Kpr*(1-NGT)
    * Little practical difference between Equations 3 and 5 unless C>>1 and NGT>>0
* Model-Fitting
    * Fit Equation 3 first to see if C>=1
    * Problems due to 5 parameters
    * Bound parameters
* Assumptions
    * ts same time each year and age
    * NGT same length each year and age

# Appendices
## Appendix 1

```{r echo=FALSE, comment=""}
cat(readLines("iCalc_tpr.R"),sep="\n")
```

## Appendix 2

```{r echo=FALSE, comment=""}
cat(readLines("VBSCGF.R"),sep="\n")
```

# References

### 2016-05-11 VBSCGF.R

@@ -2,18 +2,18 @@
 ## Main Function
 ##   Linf, t0 as usual
 ##   Kpr = K-prime as defined in Pauly et al. (units are diff than usual K)
-##   WP = "Winter Period" (middle point of no-growth period) = ts+0.5
+##   ts = start of sinusoidal growth (maximum growth rate)
 ##   NGT = "No Growth Time" = "fraction of a year where no growth occurs"
 ##   tpr = "t-prime" = actual age (t) minus cumulative NGT prior to t
 ################################################################################
 
-VBSCGF <- function(t,Linf,Kpr=NULL,t0=NULL,WP=NULL,NGT=NULL) {
+VBSCGF <- function(t,Linf,Kpr=NULL,t0=NULL,ts=NULL,NGT=NULL) {
   if (length(Linf)==5) { Kpr <- Linf[[2]]; t0 <- Linf[[3]]
-  WP <- Linf[[4]]; NGT <- Linf[[5]]
+  ts <- Linf[[4]]; NGT <- Linf[[5]]
   Linf <- Linf[[1]] }
-  tpr <- iCalc_tpr(t,WP,NGT)
+  tpr <- iCalc_tpr(t,ts,NGT)
   q <- Kpr*(tpr-t0) +
-    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(tpr-WP+1/2)) -
-    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(t0-WP+1/2))
+    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(tpr-ts)) -
+    (Kpr*(1-NGT)/(2*pi))*sin((2*pi)/(1-NGT)*(t0-ts))
   Linf*(1-exp(-q))
 }

### 2016-05-11 iCalc_tpr.R

@@ -1,12 +1,12 @@
 ################################################################################
 ## internal function to compute t-prime
 ################################################################################
-iCalc_tpr <- function(t,WP,NGT) {
+iCalc_tpr <- function(t,ts,NGT) {
   ## Step 1
-  SNG <- WP-NGT/2
+  SNG <- ts+(1-NGT)/2
   tmp.t <- t-SNG
   ## Step 2 (in parentheses) and Step 3
-  tmp.t2 <- (tmp.t-floor(tmp.t)) - NGT
+  tmp.t2 <- (tmp.t-floor(tmp.t))-NGT
   ## Step 4
   tmp.t2[tmp.t2<0] <- 0
   ## Step 5 (in parentheses) and Step 6 (also returns value)

### 2016-05-15 manuscript_v0_2.Rmd
Added references, moved real example up, some editing

@@
 author: "Derek H. Ogle"
-date: "April 29, 2016"
+date: '`r format(Sys.time(), "%A, %b %d, %Y")`'
 output:
@@
 eqncaps("PaulyMod")
-eqncaps("PaulyMod2")
 
@@
 
+----
+
+# Version notes
+
+1. Likely will not include simulated data results at the bottom.  They are a weak test at best.
+1. Would like to add 1-3 more examples of real data with varying seasonal effects (clear cessation (like Bonito), clear cessation with decline in mean length-at-age, and no clear cessation).
+1. The citation styles are not correct in this automated version.
+
+----
+
+\
+
+\
+
+\
+
 # Introduction
 
-The mean length-at-age for many animals (e.g., CITATIONS) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).
+The mean length-at-age for many fish [@Haddon2011] and other aquatic animals (e.g., @Hota1994; @Harwoodetal2014) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).
 The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as
@@
 
-Many animals exhibit seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (CITATIONS).
+Many fish exhibit seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (e.g., @Bayley1988; @Paulyetal1992; @Baconetal2005; @GarciaBerthouetal2012; @CarmonaCatotetal2014).
 `r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sine function, to model these seasonal oscillations in growth.
@@
 Note that $WP=t_{s}+\frac{1}{2}$ because the sine function in `r eqncaps("SomersMod",display="cite")` has a period (i.e., the growth period) of one year.
-Some confusion has surrounded the use of `r eqncaps("SomersMod",display="cite")`, though @GarciaBerthouetal2012 carefully clarified its form.
+Some confusion has surrounded the use of `r eqncaps("SomersMod",display="cite")`, although @GarciaBerthouetal2012 carefully clarified its form.
 
@@
 A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].
-@Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period where mean length was not allowed to decrease and that included a smooth transition of the modeled mean length-at-age into and out of the no-growth period.
+@Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period where mean length was not allowed to decrease and smoothly transitioned into and out of the no-growth period.
 Specifically, their modification is
@@
 
-@Paulyetal1992 devised  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the seasonal oscillation to the growth period and noting that $K^{'}$ only operates during the growth period).
+@Paulyetal1992 devised  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 and replacing $2\pi$ in `r eqncaps("SomersMod",display="cite")` with $\frac{2\pi}{1-NGT}$ (i.e., restricting the seasonal oscillation to the growth period and noting that $K^{'}$ only operates during the growth period).
 Their modification may be described geometrically (though not algorithmically) in two steps.
-First, the seasonal growth function with $C$=1 in `r eqncaps("SomersMod",display="cite")` is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).
-The growth trajectory is then separated at each $WP$ and horizontal segments that are $NGT$ units long are inserted at these points.
+First, `r eqncaps("SomersMod",display="cite")` with (fixed) $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).
+This growth trajectory is then separated at each $WP$ and horizontal segments that are $NGT$ units long are inserted at these points.
 This forms a growth trajectory that smoothly transitions into and out of the no-growth periods (`r figcaps("PaulyExPlot",display="cite")`).
@@
 This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.
-Thus, the values for $t^{'}$ change with each iteration of the non-linear model-fitting process.
+Thus, the values for $t^{'}$ change with each iteration of the non-linear model-fitting algorithm.
 
@@
 In practice, the calculation of $t^{'}$ also depends on the position of the no-growth period within a year.
-Here, the position of the no-growth period will be defined relative to $WP$ and $NGT$.
-The following algorithm may be used to convert from observed ages ($t$) to ages adjusted for cumulative $NGT$ prior to age $t$ ($t^{'}$).
+Here, the position of the no-growth period is defined relative to $WP$ and $NGT$, such that the following algorithm may be used to convert from observed ages ($t$) to ages adjusted for cumulative $NGT$ prior to age $t$ ($t^{'}$).
 
 1.
-Shift the age ($t$) by subtracting the start of the no-growth ($SNG$) period (i.e., $SNG=WP-\frac{NGT}{2}$) from $t$, such that a whole number will represent the start of a no-growth period.
+Shift the age ($t$) by subtracting the start of the no-growth ($SNG$) period (i.e., $SNG=WP-\frac{NGT}{2}$$=t_{s}+\frac{1}{2}-\frac{NGT}{2}$) from $t$, such that a whole number will represent the start of a no-growth period.
 For example, if $SNG$=0.4, then $t$=2.4 will become 2.0 and $t$=2.9 will become 2.5.
@@
 # Fitting the Function
+## Real Data
+
+@Stewartetal2013 examined the growth of 215 Bonito (*Sarda australis*) sampled from commercial landings.
+Fork lengths (mm) were measured for each fish and decimal ages were recorded as the number of opaque zones observed on otolith thin sections plus the proportion of the year after the designated birthdate (see @Stewartetal2013 for more detailed methods).
+@Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to not exceed 1.
+Their model fit resulted in the boundary condition of $C=1$, which suggested that Bonito ceased to grow at *at least* one point.
+This result suggests that `r eqncaps("PaulyMod",display="cite")` may fit these data better and allow the length of the no-growth period to be estimated [@Paulyetal1992].
+
+```{r echo=FALSE, results='hide', message=FALSE}
+## Bonito data
+bon <- read.csv("Bonito.csv")
+
+################################################################################
+## Purely Demonstration
+## See if results match that from Stewart et al. (2013)
+lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
+uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=1,ts=1)
+svBon <- list(Linf=72,K=0.27,t0=0,C=0.6,ts=0.2)
+fitBon1 <- nls(FL~vbSO(Age,Linf,K,t0,C,ts),data=bon,
+               start=svBon,lower=lwrbnd,upper=uprbnd,
+               algorithm="port",control=list(maxiter=100))
+## Linf=71.9, K=0.27, t0=-1.92, C=1, and ts=0.09 ... they all match (within rounding)
+cfBon1 <- cbind(Est=coef(fitBon1),confint(fitBon1))
+################################################################################
+```
+
+```{r echo=FALSE, results='hide', message=FALSE}
+## Fit new model to Bonito data
+lwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
+uprbnd <- c(Linf=Inf,Kpr=Inf,t0=Inf,ts=1,NGT=1)
+svBon <- list(Linf=72,Kpr=0.27,t0=0,ts=0.5,NGT=0.2)
+fitBon2 <- nls(FL~VBSCGF(Age,Linf,Kpr,t0,ts,NGT),data=bon,
+             start=svBon,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+cfBon2 <- cbind(Est=coef(fitBon2),confint(fitBon2))
+```
+
+`r eqncaps("PaulyMod",display="cite")` fit the Bonito data slightly better than `r eqncaps("SomersMod",display="cite")` with slightly lower residual sums-of-squares (RSS) and Akaike Information Criterion (AIC) values.
+The length of the no-growth period was estimated to be `r formatC(cfBon2["NGT","Est"],format="f",digits=3)` or `r formatC(cfBon2["NGT","Est"]*100,format="f",digits=1)`% of the year.
+The $t_{s}$ parameters were equal and the $L_\infty$ parameters were similar, but the $t_{0}$ parameters differed somewhat between the two models (`r tblcaps("BonitoRes",display="cite")`).
+The $K$ from `r eqncaps("SomersMod",display="cite")` was equal to $K'$ from `r eqncaps("PaulyMod",display="cite")` multiplied by $1-NGT$ (`r tblcaps("BonitoRes",display="cite")`).
+Graphically, there was little perceptual difference in the model fits (`r figcaps("BonitoRes",display="cite")`).
+
+```{r echo=FALSE, results="hide"}
+## A plot
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+plot(FL~Age,data=bon,pch=19,col=col2rgbt("black",1/4),
+     xlab="Age (years)",ylab="Fork Length (mm)",xlim=c(0,4),ylim=c(0,70))
+curve(VBSCGF(x,coef(fitBon2)),from=0,to=4,lwd=4,add=TRUE)
+curve(vbSO(x,coef(fitBon1)),from=0,to=4,lwd=2,add=TRUE,col="gray50")
+
+figcaps("BonitoRes",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite'),", with fixed C=1, superimposed (gray line).
+The parameter estimates (and 95% confidence intervals) from the model fits are shown in ",tblcaps("BonitoRes",display='cite'),"."))
+```
+
+`r figcaps("BonitoRes")`
+
+\
+
+```{r echo=FALSE, results="hide"}
+cfBon1 <- rbind(cfBon1,c(sum(summary(fitBon1)$residuals^2),NA,NA),c(AIC(fitBon1),NA,NA))
+cfBon2 <- rbind(cfBon2,c(sum(summary(fitBon2)$residuals^2),NA,NA),c(AIC(fitBon2),NA,NA))
+rownames(cfBon1)[6:7] <- rownames(cfBon2)[6:7] <- c("RSS","AIC")
+
+tblcaps("BonitoRes",paste0("Parameter estimates, residual sums-of-squares (RSS), and Akaike Information Criterion (AIC) from the fits of ",eqncaps('SomersMod',display='cite')," and ",eqncaps('PaulyMod',display='cite')," to the Bonito data."))
+```
+
+`r tblcaps("BonitoRes")`
+
+```{r echo=FALSE, comment=""}
+cat("Somers Model (Equation 3) fit")
+print(round(cfBon1,2),na.print="-")
+cat("Modified Pauly Model (Equation 4) fit")
+print(round(cfBon2,2),na.print="-")
+```
+
 ## Simulated Data
@@
 One example, for a "long" $NGT$, "late" $t_{s}$, and low individual variability, is shown in `r figcaps("PaulySimRes",display="cite")`.
-The results from fitting `r eqncaps("PaulyMod",display="cite")` to these data sets are not surprising as one would expect the parameter estimates to be very close to the values used to create them, given that `r eqncaps("PaulyMod2",display="cite")` was used to both create and fit the data.
+The results from fitting `r eqncaps("PaulyMod",display="cite")` to these data sets are not surprising as one would expect the parameter estimates to be very close to the values used to create them, given that `r eqncaps("PaulyMod",display="cite")` was used to both create and fit the data.
 
@@
 
-tblcaps("PaulySimRes",paste0("Parameter estimates from fitting ",eqncaps('PaulyMod2',display='cite')," to simulated data sets using $L_{\\infty}$=",Linf,", $K'$=",Kpr,", $t_{0}$=",t0,", and varying values of $t_{s}$, $NGT$, and $\\sigma$."))
+tblcaps("PaulySimRes",paste0("Parameter estimates from fitting ",eqncaps('PaulyMod',display='cite')," to simulated data sets using $L_{\\infty}$=",Linf,", $K'$=",Kpr,", $t_{0}$=",t0,", and varying values of $t_{s}$, $NGT$, and $\\sigma$."))
 ```
@@
 \
-
-## Real Data
-
-@Stewartetal2013 examined the growth of 215 Bonito (*Sarda australis*) sampled from commercial landings.
-Fork lengths (mm) were measured for each fish and decimal ages were recorded as the number of opaque zones observed on otolith thin sections plus the proportion of the year after the designated birthdate (see @Stewartetal2013 for more detailed methods).
-@Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to not exceed 1.
-Their model fit resulted in the boundary condition of $C=1$, which suggested that Bonito ceased to grow at *at least* one point.
-This result suggests that `r eqncaps("PaulyMod2",display="cite")` should be fit to these data to determine the length of the no-growth period [@Paulyetal1992].
-
-```{r echo=FALSE, results='hide', message=FALSE}
-## Bonito data
-bon <- read.csv("Bonito.csv")
-
-################################################################################
-## Purely Demonstration
-## See if results match that from Stewart et al. (2013)
-lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
-uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=1,ts=1)
-svBon <- list(Linf=72,K=0.27,t0=0,C=0.6,ts=0.2)
-fitBon1 <- nls(FL~vbSO(Age,Linf,K,t0,C,ts),data=bon,
-               start=svBon,lower=lwrbnd,upper=uprbnd,
-               algorithm="port",control=list(maxiter=100))
-## Linf=71.9, K=0.27, t0=-1.92, C=1, and ts=0.09 ... they all match (within rounding)
-cfBon1 <- cbind(Est=coef(fitBon1),confint(fitBon1))
-################################################################################
-```
-
-```{r echo=FALSE, results='hide', message=FALSE}
-## Fit new model to Bonito data
-lwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
-uprbnd <- c(Linf=Inf,Kpr=Inf,t0=Inf,ts=1,NGT=1)
-svBon <- list(Linf=72,Kpr=0.27,t0=0,ts=0.5,NGT=0.2)
-fitBon2 <- nls(FL~VBSCGF(Age,Linf,Kpr,t0,ts,NGT),data=bon,
-             start=svBon,lower=lwrbnd,upper=uprbnd,
-             algorithm="port",control=list(maxiter=100))
-cfBon2 <- cbind(Est=coef(fitBon2),confint(fitBon2))
-```
-
-`r eqncaps("PaulyMod",display="cite")` fit the Bonito data slightly better than `r eqncaps("SomersMod",display="cite")` with slightly lower residual sums-of-squares (RSS) and Akaike Information Criterion (AIC) values.
-The length of the no-growth period was estimated to be `r formatC(cfBon2["NGT","Est"],format="f",digits=3)` or `r formatC(cfBon2["NGT","Est"]*100,format="f",digits=1)`% of the year.
-The $t_{s}$ parameters were equal, the $L_\infty$ parameters were similar, but the $t_{0}$ parameters differed somewhat between the two models.
-Finally, $K$ from `r eqncaps("SomersMod",display="cite")` was equal to $K'$ from `r eqncaps("PaulyMod2",display="cite")` multiplied by $1-NGT$.
-
-```{r echo=FALSE, results="hide"}
-## A plot
-par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
-plot(FL~Age,data=bon,pch=19,col="black",
-     xlab="Age (years)",ylab="Fork Length (mm)",xlim=c(0,4),ylim=c(0,70))
-curve(VBSCGF(x,coef(fitBon2)),from=0,to=4,lwd=4,add=TRUE)
-curve(vbSO(x,coef(fitBon1)),from=0,to=4,lwd=2,add=TRUE,col="gray50")
-
-figcaps("BonitoRes",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod2',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite')," (gray line) superimposed.
-The parameter estimates (and 95% confidence intervals) from the model fits are shown in ",tblcaps("BonitoRes",display='cite'),"."))
-```
-
-`r figcaps("BonitoRes")`
-
-\
-
-```{r echo=FALSE, results="hide"}
-cfBon1 <- rbind(cfBon1,c(sum(summary(fitBon1)$residuals^2),NA,NA),c(AIC(fitBon1),NA,NA))
-cfBon2 <- rbind(cfBon2,c(sum(summary(fitBon2)$residuals^2),NA,NA),c(AIC(fitBon2),NA,NA))
-rownames(cfBon1)[6:7] <- rownames(cfBon2)[6:7] <- c("RSS","AIC")
-
-tblcaps("BonitoRes",paste0("Parameter estimates, residual sums-of-squares (RSS), and Akaike Information Criterion (AIC) from the fits of ",eqncaps('SomersMod',display='cite')," and ",eqncaps('PaulyMod2',display='cite')," to the Bonito data."))
-```
-
-`r tblcaps("BonitoRes")`
-
-```{r echo=FALSE, comment=""}
-cat("Somers Model (Equation 3) fit")
-print(round(cfBon1,2),na.print="-")
-cat("Modified Pauly Model (Equation 6) fit")
-print(round(cfBon2,2),na.print="-")
-```
 
@@
 
+# Acknowledgments
+
+John Stewart (New South Wales Department of Primary Industries Fisheries) graciously provided the Bonito length-at-age data.
+
 # Appendices

### 2016-05-22 zzzOld/doc/manuscript_v0.Rmd
Moved to Word, Reorganized file strux, added mosquitofish example

--- 2016-05-09 manuscript_v0_1.Rmd
@@
 A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].
-@Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period where mean length was not allowed to decrease and that included a smooth transition of the modeled mean length-at-age into and out of the no-growth period.
+To address this unrealistic decrease in mean length, @Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period that included a smooth transition of the modeled mean length-at-age into and out of the no-growth period.
 Specifically, their modification is
@@
 
-\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$" [@Paulyetal1992].
-Furthermore,  @Paulyetal1992 noted that the units of $K$ changed from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")`.
-To eliminate confusion, they suggested using $K^{'}$ in `r eqncaps("PaulyMod",display="cite")`, as we do here.
+\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$."  Furthermore,  @Paulyetal1992 noted that the units of $K$ changed from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")`.
+To eliminate confusion, he suggested using $K^{'}$ in `r eqncaps("PaulyMod",display="cite")`, as we do here.
 
-@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth and noting that $K^{'}$ only operates during this shorter period).
+@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth and noting that $K^{'}$ only operates in this shorter period).
 Their modification may be described geometrically (though not algorithmically) in two steps.
@@
 
-# The Modified Seasonal Cessation Growth Function
+# The Modified Model
 
@@
 q= & K^{'}(t^{'}-t_{0}) \\
-   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t^{'}-WP-\frac{1}{2}\right)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`} \\
-   & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t_{0}-WP-\frac{1}{2}\right)\right)
+   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t^{'}-WP-\frac{1}{2}\right)\right)-\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t_{0}-WP-\frac{1}{2}\right)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`}
 \end{aligned}  
@@
 In practice, the calculation of $t^{'}$ also depends on the position of the no-growth period within a year.
-Here, the position of the no-growth period is defined by $SNG$ and $NGT$.
-However, because the $SNG$ has a one-to-one relation with $WP$ and $t_{s}$, the no-growth period can also be defined relative to these parameters.
+Here, the position of the no-growth period is defined by $SNG$.
+However, because the $SNG$ has a one-to-one relation with $WP$ and $t_{s}$ it can also be defined relative to these parameters.
 
@@
 # Fitting the Modified Model
-Simple simulated data were initially used to demonstrate the fitting of `r eqncaps("paulyMod2",display="cite")`.
+Simple simulated data were used to demonstrate the fitting of `r eqncaps("paulyMod2",display="cite")`.
 The data were simulated by randomly selecting 200 real numbers from a uniform distribution between 0 and 5 to serve as observed ages, plugging these ages into `r eqncaps("paulyMod2",display="cite")` to compute a mean length for each age, and then adding a random deviate from a normal distribution with a mean of 0 and a standard deviation of $\sigma$ to each mean length to simulate observed individual lengths.
@@
 The "port" algorithm in `nls()` was used as $L_{\infty}$ and $K^{'}$ were constrained to be positive and $WP$ and $NGT$ were constrained to be between 0 and 1.
-One would expect the parameter estimates to be very close to the values used to create the simulated data, as `r eqncaps("paulyMod2",display="cite")` was used to both create and fit the data.
+One would expect the parameter estimates to be very close to the values used to create the simulated data as `r eqncaps("paulyMod2",display="cite")` was used to both create and fit the data.
 
 One example of a model fit to data that were simulated to have a "long" NGT, "late" WP, and low individual variability is shown in `r figcaps("PaulyExData1",display="cite")`.
-In this example, the parameter estimates are quite close to the parameter values used to simulate the data (the set parameter values are within the confidence intervals for each parameter).
+In this example, the parameter estimates are quite close to the parameter values used to simulate the data (the set parameter values are within the confidence interval for each parameter).
 
@@
 figcaps("PaulyExData1",paste0("Example VBGF using ",eqncaps('PaulyMod2',display='cite')," with parameters shown in the upper-left corner and $\\sigma$=",sigma,".
-Parameter estimates (and 95% confidence intervals) from non-linear regression are shown the lower-right corner.
+Paramter estimates (and 95% confidence intervals) from non-linear regression are shown the lower-right corner.
 Each WP is shown by the arrow and each NGT is shown by the horizontal interval bar centered on the WP arrow."))
@@
 `r figcaps("PaulyExData1")`
-
-\
-
-`r eqncaps("paulyMod2",display="cite")` was fit to simulated data sets that consisted of all combinations of "early" ($WP=0.2$) and "late" ($WP=0.8$) winter points, "short" ($NGT=0.2$) and "long" ($NGT=0.5$) no-growth periods, and "low" ($\sigma=0.3$) and "high" ($\sigma=0.5$) individual variabilities.
-These limited simulations suggest that all parameters are consistently well estimated, with the possible exception of $t_{0}$.
 
@@
 
+```{r echo=FALSE, results='hide',eval=FALSE}
+## Set some common parameters
+Linf <- 30; K <- 0.3; t0 <- -0.1
+## Set bounds for model-fitting
+lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
+uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
+```
+
+```{r echo=FALSE, results="hide", message=FALSE,eval=FALSE}
+## Setup
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+## Set some specific growth parameters
+# fairly typical north temperate situation
+WP <- 0.8; NGT <- 0.5
+gper <- c(WP+NGT/2-1,WP-NGT/2)*365
+## Actually generate the data (with very little variability)
+d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.001,SD=0.001,paramCV=0.001,
+                growth.per=gper,sample.per=c(1,365),lendigs=1)
+## Try fitting the VBSCGF model
+sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
+fitP1 <- nls(lenCap~VBSCGF(ageFracY,Linf,K,t0,WP,NGT),data=d1,
+             start=sv,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+( cfs <- coef(fitP1) )
+## A plot
+plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/3),
+     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
+curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
+WPs <- 1:5+cfs[["WP"]]
+LatWPs <- VBSCGF(WPs,Linf=cfs)
+arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
+SNGTs <- WPs-cfs[["NGT"]]/2
+ENGTs <- WPs+cfs[["NGT"]]/2
+arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)
+
+tmp <- c("Linf=","K=","t0=","WP=","NGT=")
+legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP,NGT)),
+       bty="n",inset=-0.02,cex=0.8)
+cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
+legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
+```
+
+```{r echo=FALSE, results="hide", message=FALSE,eval=FALSE}
+## Setup
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+## Set some specific growth parameters
+# fairly typical north temperate situation
+WP <- 0.8; NGT <- 0.5
+gper <- c(WP+NGT/2-1,WP-NGT/2)*365
+## Actually generate the data (with more variability)
+d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.2,SD=0.5,paramCV=0.01,
+                growth.per=gper,sample.per=c(1,365),lendigs=1)
+## Try fitting the VBSCGF model
+sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
+fitP1 <- nls(lenCap~VBSCGF(ageFracY,Linf,K,t0,WP,NGT),data=d1,
+             start=sv,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+( cfs <- coef(fitP1) )
+## A plot
+plot(lenCap~ageFracY,data=d1,pch=19,col=col2rgbt("black",1/3),
+     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
+curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
+WPs <- 1:5+cfs[["WP"]]
+LatWPs <- VBSCGF(WPs,Linf=cfs)
+arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
+SNGTs <- WPs-cfs[["NGT"]]/2
+ENGTs <- WPs+cfs[["NGT"]]/2
+arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)
+
+tmp <- c("Linf=","K=","t0=","WP=","NGT=")
+legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP,NGT)),
+       bty="n",inset=-0.02,cex=0.8)
+cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
+legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
+```
+
+```{r echo=FALSE, results="hide", message=FALSE,eval=FALSE}
+## Setup
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
+## Set some specific growth parameters
+# fairly typical north temperate situation
+WP <- 0.8; NGT <- 0.5
+gper <- c(WP+NGT/2-1,WP-NGT/2)*365
+## Actually generate the data (with more variability)
+d1 <- vbDataGen(200,Linf=Linf,K=K,t0=t0,maxAge=5,SE=0.2,SD=0.5,paramCV=0.01,
+                growth.per=gper,sample.per=c(1,365),lendigs=1)
+## Try fitting the VBSCGF model
+sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
+fitP1 <- nls(lenCap~VBSCGF(ageFracY-0.3,Linf,K,t0,WP,NGT),data=d1,
+             start=sv,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+( cfs <- coef(fitP1) )
+## A plot
+plot(lenCap~I(ageFracY-0.3),data=d1,pch=19,col=col2rgbt("black",1/3),
+     xlab="Age (years)",ylab="Length",xlim=c(1,6),ylim=c(7,27))
+curve(VBSCGF(x,coef(fitP1)),from=1,to=6,lwd=2,add=TRUE)
+WPs <- 1:5+cfs[["WP"]]
+LatWPs <- VBSCGF(WPs,Linf=cfs)
+arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
+SNGTs <- WPs-cfs[["NGT"]]/2
+ENGTs <- WPs+cfs[["NGT"]]/2
+arrows(SNGTs,LatWPs-0.7,ENGTs,LatWPs-0.7,lwd=2,length=0.05,angle=90,code=3)
+
+tmp <- c("Linf=","K=","t0=","WP=","NGT=")
+legend("topleft",paste0(c("Set Values",tmp),c("",Linf,K,t0,WP-0.3,NGT)),
+       bty="n",inset=-0.02,cex=0.8)
+cfs <- round(cbind(Est=coef(fitP1),confint(fitP1)),3)
+legend("bottomright",c("Estimated Values",paste0(tmp,cfs[,1]," (",cfs[,2],",",cfs[,3],")")),bty="n",cex=0.8)
+```
+
 ## Real Data
 
-@Stewartetal2013 examined the growth of 215 Bonito (*Sarda australis*) sampled from commercial landings.
-Detailed methods are described in @Stewartetal2013, but note here that fork lengths (mm) were measured for each fish and ages were the decimal age calculated as the number of opaque zones observed on otolith thin sections plus the proportion of the year after the designated birthdate.
-@Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to be less than 1.
-Their model fit resulted in the boundary condition of $C=1$, which suggested that Bonito ceased to grow at *at least* one point.
-As $C$ is at least equal to 1, this result also suggests that `r eqncaps("paulyMod2",display="cite")` should be fit to these data to determine the length of the no-growth period [@Paulyetal1992].
-Thus, `r eqncaps("paulyMod2",display="cite")` is fit to these data here as an example of fitting  `r eqncaps("paulyMod2",display="cite")` to real data.
-
-```{r echo=FALSE, results='hide'}
-## Bonito data
-bon <- read.csv("Bonito.csv")
-
-################################################################################
-## Purely Demonstration
-## See if results match that from Stewart et al. (2013)
-vbSO <- vbFuns("Somers")
-lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
-uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=1,ts=1)
-svBon <- list(Linf=72,K=0.27,t0=0,C=0.6,ts=0.2)
-fitBon1 <- nls(FL~vbSO(Age,Linf,K,t0,C,ts),data=bon,
-               start=svBon,lower=lwrbnd,upper=uprbnd,
-               algorithm="port",control=list(maxiter=100))
-## Linf=71.9, K=0.27, t0=-1.92, C=1, and to=0.09 ... they all match (within rounding)
-coef(fitBon1)
-################################################################################
-```
-
-```{r echo=FALSE, results='hide', message=FALSE}
-## Fit new model to Bonito data
-lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
-uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
-svBon <- list(Linf=72,K=0.27,t0=0,WP=0.5,NGT=0.2)
-fitBon2 <- nls(FL~VBSCGF(Age,Linf,K,t0,WP,NGT),data=bon,
-             start=svBon,lower=lwrbnd,upper=uprbnd,
-             algorithm="port",control=list(maxiter=100))
-cfBon2 <- cbind(Est=coef(fitBon2),confint(fitBon2))
-```
-
-```{r echo=FALSE, results="hide"}
-## A plot
-par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
-plot(FL~Age,data=bon,pch=19,col="black",
-     xlab="Age (years)",ylab="Length",xlim=c(0,4),ylim=c(0,70))
-curve(VBSCGF(x,coef(fitBon2)),from=0,to=4,lwd=4,add=TRUE)
-curve(vbSO(x,coef(fitBon1)),from=0,to=4,lwd=2,add=TRUE,col="gray50")
-legend("bottomright",c(TeX("\\mathbf{Estimated Values}"),
-                       TeX(paste0("$L_{\\infty}$"," = ",
-                                  formatC(cfBon2["Linf","Est"],format="f",digits=1),
-                                  " (",formatC(cfBon2["Linf","2.5%"],format="f",digits=1),
-                                  ", ",formatC(cfBon2["Linf","97.5%"],format="f",digits=1),")")),
-                   paste0("K' = ",formatC(cfBon2["K","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["K","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["K","97.5%"],format="f",digits=2),")"),
-                   TeX(paste0("$t_{0}$"," = ",
-                                  formatC(cfBon2["t0","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["t0","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["t0","97.5%"],format="f",digits=2),")")),
-                   paste0("WP = ",formatC(cfBon2["WP","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["WP","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["WP","97.5%"],format="f",digits=2),")"),
-                   paste0("NGT = ",formatC(cfBon2["NGT","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["NGT","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["NGT","97.5%"],format="f",digits=2),")")),
-        bty="n",cex=0.8)
-
-figcaps("PaulyBonito",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod2',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite')," (gray line) superimposed.
-The parameter estimates (and 95% confidence intervals) from non-linear regression for ",eqncaps('PaulyMod2',display='cite')," are shown the lower-right corner."))
-```
-
-`r figcaps("PaulyBonito")`
-
-\
-
 # Discussion
-
-* General -- fits perfectly simulated data well.
-* Assumptions -- WP same time each year and age, NGT same length each year and age
-* Model-Fitting -- bounding parameters, more difficult to fit five parameters
-* Practical Difference -- not large unless C>>1 and NGT>>0
 

### 2016-05-22 zzzOld/doc/manuscript_v0_1.Rmd
Moved to Word, Reorganized file strux, added mosquitofish example

--- 2016-05-09 manuscript_v0_1.Rmd
@@
 `r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sine function, to model these seasonal oscillations in growth.
-The most popular of these modifications is from @HoenigChoudarayHanumara1982 and @Somers1988 (and carefully reiterated in @GarciaBerthouetal2012), and uses
+The most popular of these modifications, from @HoenigChoudarayHanumara1982 and @Somers1988, is
 
@@
 If $C$=1, then growth completely stops once a year at the "winter-point" ($WP$), whereas values of 0&lt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).
+Some confusion has surrounded this function, though @GarciaBerthouetal2012 carefully clarified its form.
 
@@
 
-@Paulyetal1992 derived  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the period of growth and noting that $K^{'}$ only operates during this shorter period).
+@Paulyetal1992 devised  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the growth period and noting that $K^{'}$ only operates during the growth period).
 Their modification may be described geometrically (though not algorithmically) in two steps.
-First, the seasonal growth function in `r eqncaps("SomersMod",display="cite")` with $C$=1 is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).
+First, the seasonal growth function from with $C$=1 in `r eqncaps("SomersMod",display="cite")` is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).
 The growth trajectory is then separated at each $WP$ and horizontal segments that are $NGT$ units long are inserted at these points.
@@
 WP <- 0.55; NGT <- 0.3
-Linf <- 30; K <- 0.3; Kpr <- K/(1-NGT); t0 <- 0
+Linf <- 30; Kpr <- 0.4; t0 <- 0
 Pcf <- c(Linf,Kpr,t0,WP,NGT)
@@
 PL <- VBSCGF(t,Linf=Pcf)
-## Calculate WPs and start and end of NGT
+## Calculate WPs and start and end of each NGT
 # find WPs and Lengths at WPs
-WPs <- 0:2+WP
-LatWPs <- VBSCGF(WPs,Linf=Pcf)
 # find start and end of NGT
-NGTs <- WPs-NGT/2
-NGTe <- WPs+NGT/2
 ## Make schematic plot
@@
 ## Put on NGT polygons
-for (i in 1:length(NGTs)) polygon(c(NGTs[i],NGTs[i],NGTe[i],NGTe[i]),c(0,22,22,0),
+WPs <- 0:2+WP
+LatWPs <- VBSCGF(WPs,Linf=Pcf)
+SNGTs <- WPs-NGT/2
+ENGTs <- WPs+NGT/2
+for (i in 1:length(SNGTs)) polygon(c(SNGTs[i],SNGTs[i],ENGTs[i],ENGTs[i]),c(0,LatWPs[i],LatWPs[i],0),
                                    col=col2rgbt("black",1/20),border=NA)
+## Mark the WPs
+arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
+## Mark the NGTs
+arrows(SNGTs,LatWPs+1.2,ENGTs,LatWPs+1.2,lwd=2,length=0.025,angle=90,code=3)
 ## Put on the growth trajectory
 lines(PL~t,lwd=2)
-## Place a point at the WP
-points(WPs,LatWPs,pch=3,col="gray50")
 ## Add some important values to the age axis
 axis(1,0:3)
-axis(1,NGTs,tcl=-0.2)
-axis(1,NGTe,tcl=-0.2)
+axis(1,SNGTs,tcl=-0.2)
+axis(1,ENGTs,tcl=-0.2)
 ## Add the t-prime labels
 # makes inside ticks
-axis(1,at=c(0:3,NGTs,NGTe),labels=NA,tcl=0.2)
+axis(1,at=c(0:3,SNGTs,ENGTs),labels=NA,tcl=0.2)
 axis(1,at=1:3,labels=iCalc_tpr(1:3,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
-axis(1,at=NGTs,labels=iCalc_tpr(NGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
-axis(1,at=NGTe,labels=iCalc_tpr(NGTe,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=SNGTs,labels=iCalc_tpr(SNGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
+axis(1,at=ENGTs,labels=iCalc_tpr(ENGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
 axis(1,at=3.3,labels="<= t",tick=FALSE)
@@
 figcaps("PaulyExPlot",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," or ",eqncaps('PaulyMod2',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",formatC(Kpr,format="f",digits=4),", $t_{0}$=",t0,", $NGT$=",NGT,", and $t_{s}$=",WP-0.5," or $WP$=",WP,".
-The no-growth periods are marked with gray bars and each $WP$ is marked with a gray plus symbol.
+Each $WP$ is shown by the arrow and each no-growth period is shown by the horizontal error bar centered on the $WP$ arrow and by the gray region that extends to the x-axis.
 The ages adjusted for the $NGT$ (i.e., $t^{'}$) are shown above the x-axis."))
@@
 This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.
-In other words, the values for $t^{'}$ change with each iteration of the non-linear model-fitting process.
+Thus, the values for $t^{'}$ change with each iteration of the non-linear model-fitting process.
 
-Therefore, the objectives of this note are to (i) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while providing a more meaningful parameter; (ii) describe a function for calculating $t^{'}$ that may be used in the model-fitting process; (iii) provide an (open-source) algorithm for the modified function that can be used in model-fitting; and (iv) demonstrate the use of the modified function for fitting length-at-age data.
+Therefore, the objectives of this note are to (i) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while also providing a more meaningful parameter; (ii) operationalize the calculation of $t^{'}$ for use in model fitting; (iii) provide an (open-source) algorithm for the modified function for use in model fitting; and (iv) demonstrate the fitting of the modified function to length-at-age data.
 
@@
 
-In `r eqncaps("SomersMod",display="cite")`, $WP=t_{s}+\frac{1}{2}$ because the sine function has a period of one year.
-The growth period is compressed in `r eqncaps("PaulyMod",display="cite")` to be $1-NGT$.
-Thus, the start of the no-growth period ($SNG$) is $t_{s}+\frac{1-NGT}{2}$.
-The center of the no-growth period is then $SNG+\frac{NGT}{2}$ or $t_{s}+\frac{1}{2}$.
-Thus, $WP=t_{s}+\frac{1}{2}$ is the center of the no-growth period for `r eqncaps("PaulyMod",display="cite")`.
+In `r eqncaps("SomersMod",display="cite")`, $WP=t_{s}+\frac{1}{2}$ because the sine function has a period (i.e., the growth period) of one year.
+Because the growth period is compressed in `r eqncaps("PaulyMod",display="cite")` to be $1-NGT$, the start of the no-growth period in `r eqncaps("PaulyMod",display="cite")` is $t_{s}+\frac{1-NGT}{2}$.
+The center of the no-growth period is the start of the no-growth period plus half of the no-growth time, or
+
+$$ t_{s}+\frac{1-NGT}{2}+\frac{NGT}{2} = t_{s}+\frac{1}{2} = WP $$.
+
+Thus, $WP=t_{s}+\frac{1}{2}$ is the center of the no-growth period in `r eqncaps("PaulyMod",display="cite")`.
 
@@
 This algorithm for computing $t^{'}$ is implemented in an R [@R2016] function as shown in Appendix 1.
-With this, `r eqncaps("paulyMod2",display="cite")` is easly implemented as an R function as shown in Appendix 2.
+With this, `r eqncaps("PaulyMod2",display="cite")` is easly implemented as an R function as shown in Appendix 2.
 
 # Fitting the Modified Model
-Simple simulated data were initially used to demonstrate the fitting of `r eqncaps("paulyMod2",display="cite")`.
-The data were simulated by randomly selecting 200 real numbers from a uniform distribution between 0 and 5 to serve as observed ages, plugging these ages into `r eqncaps("paulyMod2",display="cite")` to compute a mean length for each age, and then adding a random deviate from a normal distribution with a mean of 0 and a standard deviation of $\sigma$ to each mean length to simulate observed individual lengths.
-The `nls()` function from R was then used to estimate parameter values for `r eqncaps("paulyMod2",display="cite")` from the simulated data.
+Simple simulated data were initially used to demonstrate the fitting of `r eqncaps("PaulyMod2",display="cite")`.
+The data were simulated by randomly selecting 200 real numbers from a uniform distribution between 0 and 5 to serve as observed ages, plugging these ages into `r eqncaps("PaulyMod2",display="cite")` to compute a mean length for each age, and then adding a random deviate from a normal distribution with a mean of 0 and a standard deviation of $\sigma$ to each mean length to simulate observed individual lengths.
+The `nls()` function from R was then used to estimate parameter values for `r eqncaps("PaulyMod2",display="cite")` from the simulated data.
 The "port" algorithm in `nls()` was used as $L_{\infty}$ and $K^{'}$ were constrained to be positive and $WP$ and $NGT$ were constrained to be between 0 and 1.
-One would expect the parameter estimates to be very close to the values used to create the simulated data, as `r eqncaps("paulyMod2",display="cite")` was used to both create and fit the data.
+One would expect the parameter estimates to be very close to the values used to create the simulated data, as `r eqncaps("PaulyMod2",display="cite")` was used to both create and fit the data.
 
-One example of a model fit to data that were simulated to have a "long" NGT, "late" WP, and low individual variability is shown in `r figcaps("PaulyExData1",display="cite")`.
+One example of a model fit to data that were simulated to have a "long" NGT, "late" WP, and low individual variability is shown in `r figcaps("PaulySimRes",display="cite")`.
 In this example, the parameter estimates are quite close to the parameter values used to simulate the data (the set parameter values are within the confidence intervals for each parameter).
@@
 
-figcaps("PaulyExData1",paste0("Example VBGF using ",eqncaps('PaulyMod2',display='cite')," with parameters shown in the upper-left corner and $\\sigma$=",sigma,".
+figcaps("PaulySimRes",paste0("Example VBGF using ",eqncaps('PaulyMod2',display='cite')," with parameters shown in the upper-left corner and $\\sigma$=",sigma,".
 Parameter estimates (and 95% confidence intervals) from non-linear regression are shown the lower-right corner.
@@
 
-`r figcaps("PaulyExData1")`
+`r figcaps("PaulySimRes")`
 
@@
 
-`r eqncaps("paulyMod2",display="cite")` was fit to simulated data sets that consisted of all combinations of "early" ($WP=0.2$) and "late" ($WP=0.8$) winter points, "short" ($NGT=0.2$) and "long" ($NGT=0.5$) no-growth periods, and "low" ($\sigma=0.3$) and "high" ($\sigma=0.5$) individual variabilities.
+`r eqncaps("PaulyMod2",display="cite")` was fit to simulated data sets that consisted of all combinations of "early" ($WP=0.2$) and "late" ($WP=0.8$) winter points, "short" ($NGT=0.2$) and "long" ($NGT=0.5$) no-growth periods, and "low" ($\sigma=0.3$) and "high" ($\sigma=0.5$) individual variabilities.
 These limited simulations suggest that all parameters are consistently well estimated, with the possible exception of $t_{0}$.
@@
 
-```{r echo=FALSE, message=FALSE}
+```{r echo=FALSE, results="hide", message=FALSE}
 WPs <- rep(c(0.8,0.2),each=4)
@@
 colnames(res) <- c("WPset","NGTset","sigmaset","WPest","WPlci","WPuci","NGTest","NGTlci","NGTuci","Linfest","Linflci","Linfuci","Kest","Klci","Kuci","t0est","t0lci","t0uci")
+
+tblcaps("PaulySimRes",paste0("Parameter estimates from ",eqncaps('SomersMod',display='cite')," for simulated data using $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", and varying values of $WP$, $NGT$, and $\\sigma$."))
+```
+
+`r tblcaps("PaulySimRes")`
+
+```{r echo=FALSE, comment=""}
 res
 ```
+
+\
 
@@
 Detailed methods are described in @Stewartetal2013, but note here that fork lengths (mm) were measured for each fish and ages were the decimal age calculated as the number of opaque zones observed on otolith thin sections plus the proportion of the year after the designated birthdate.
-@Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to be less than 1.
+@Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to not exceed 1.
 Their model fit resulted in the boundary condition of $C=1$, which suggested that Bonito ceased to grow at *at least* one point.
-As $C$ is at least equal to 1, this result also suggests that `r eqncaps("paulyMod2",display="cite")` should be fit to these data to determine the length of the no-growth period [@Paulyetal1992].
-Thus, `r eqncaps("paulyMod2",display="cite")` is fit to these data here as an example of fitting  `r eqncaps("paulyMod2",display="cite")` to real data.
+As $C$ is at least equal to 1, this result also suggests that `r eqncaps("PaulyMod2",display="cite")` should be fit to these data to determine the length of the no-growth period [@Paulyetal1992].
+Thus, `r eqncaps("PaulyMod2",display="cite")` is fit to these data here as an example.
 
-```{r echo=FALSE, results='hide'}
+```{r echo=FALSE, results='hide', message=FALSE}
 ## Bonito data
@@
                algorithm="port",control=list(maxiter=100))
-## Linf=71.9, K=0.27, t0=-1.92, C=1, and to=0.09 ... they all match (within rounding)
-coef(fitBon1)
+## Linf=71.9, K=0.27, t0=-1.92, C=1, and ts=0.09 ... they all match (within rounding)
+cfBon1 <- cbind(Est=coef(fitBon1),confint(fitBon1))
 ################################################################################
@@
 
+`r eqncaps("PaulyMod2",display="cite")` fit the Bonito data slightly better than `r eqncaps("SomersMod",display="cite")` with slightly lower residual sums-of-squares (RSS) and Akaike Information Criterion (AIC) values.
+The length of the no-growth period was estimated to be `r formatC(cfBon2["NGT","Est"],format="f",digits=3)` or `r formatC(cfBon2["NGT","Est"]*100,format="f",digits=1)`% of the year.
+The $WP$ parameter from `r eqncaps("PaulyMod2",display="cite")` was equal to the $t_{s}$ parameter from `r eqncaps("SomersMod",display="cite")` plus 0.5.
+The $L_\infty$ parameters were very similar between the two models, but the $t_{0}$ parameters differed somewhat.
+Finally, $K$ from `r eqncaps("SomersMod",display="cite")` was equal to $K'$ from `r eqncaps("PaulyMod2",display="cite")` divided by $1-NGT$.
+
 ```{r echo=FALSE, results="hide"}
@@
 plot(FL~Age,data=bon,pch=19,col="black",
-     xlab="Age (years)",ylab="Length",xlim=c(0,4),ylim=c(0,70))
+     xlab="Age (years)",ylab="Fork Length (mm)",xlim=c(0,4),ylim=c(0,70))
 curve(VBSCGF(x,coef(fitBon2)),from=0,to=4,lwd=4,add=TRUE)
 curve(vbSO(x,coef(fitBon1)),from=0,to=4,lwd=2,add=TRUE,col="gray50")
-legend("bottomright",c(TeX("\\mathbf{Estimated Values}"),
-                       TeX(paste0("$L_{\\infty}$"," = ",
-                                  formatC(cfBon2["Linf","Est"],format="f",digits=1),
-                                  " (",formatC(cfBon2["Linf","2.5%"],format="f",digits=1),
-                                  ", ",formatC(cfBon2["Linf","97.5%"],format="f",digits=1),")")),
-                   paste0("K' = ",formatC(cfBon2["K","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["K","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["K","97.5%"],format="f",digits=2),")"),
-                   TeX(paste0("$t_{0}$"," = ",
-                                  formatC(cfBon2["t0","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["t0","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["t0","97.5%"],format="f",digits=2),")")),
-                   paste0("WP = ",formatC(cfBon2["WP","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["WP","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["WP","97.5%"],format="f",digits=2),")"),
-                   paste0("NGT = ",formatC(cfBon2["NGT","Est"],format="f",digits=2),
-                                  " (",formatC(cfBon2["NGT","2.5%"],format="f",digits=2),
-                                  ", ",formatC(cfBon2["NGT","97.5%"],format="f",digits=2),")")),
-        bty="n",cex=0.8)
 
-figcaps("PaulyBonito",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod2',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite')," (gray line) superimposed.
+figcaps("BonitoRes",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod2',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite')," (gray line) superimposed.
 The parameter estimates (and 95% confidence intervals) from non-linear regression for ",eqncaps('PaulyMod2',display='cite')," are shown the lower-right corner."))
@@
 
-`r figcaps("PaulyBonito")`
+`r figcaps("BonitoRes")`
+
+\
+
+```{r echo=FALSE, results="hide"}
+cfBon1 <- rbind(cfBon1,c(sum(summary(fitBon1)$residuals^2),NA,NA),c(AIC(fitBon1),NA,NA))
+cfBon2 <- rbind(cfBon2,c(sum(summary(fitBon2)$residuals^2),NA,NA),c(AIC(fitBon2),NA,NA))
+rownames(cfBon1)[6:7] <- rownames(cfBon2)[6:7] <- c("RSS","AIC")
+
+tblcaps("BonitoRes",paste0("Parameter estimates, residual sums-of-squares (RSS), and Akaike Information Criterion (AIC) from the fits of ",eqncaps('SomersMod',display='cite')," and ",eqncaps('PaulyMod2',display='cite')," to the Bonito data."))
+```
+
+`r tblcaps("BonitoRes")`
+
+```{r echo=FALSE, comment=""}
+cat("Somers Model (Equation 3) fit")
+print(round(cfBon1,2),na.print="-")
+cat("Modified Pauly Model (Equation 6) fit")
+print(round(cfBon2,2),na.print="-")
+```
 

### 2016-05-22 zzzOld/doc/manuscript_v0_2.Rmd
Moved to Word, Reorganized file strux, added mosquitofish example

--- 2016-05-15 manuscript_v0_2.Rmd
@@
 ---
-title: "Revisiting the von Bertalanffy Seasonal Cessational Growth Function of Pauly et al. (1992)"
+title: Revisiting the von Bertalanffy Seasonal Cessational Growth Function of Pauly
+  et al. (1992)
 author: "Derek H. Ogle"
@@
 output:
+  word_document: 
+    fig_height: 3.5
+    fig_width: 4.5
   pdf_document:

### 2016-05-22 zzzOld/doc/manuscript_v0_3.Rmd
Moved to Word, Reorganized file strux, added mosquitofish example

--- 2016-05-22 zzzOld/doc/manuscript_v0_2.Rmd
@@
 # Fitting the Function
-## Real Data
+## Bonito Data
 
@@
 print(round(cfBon2,2),na.print="-")
+```
+
+## Mosquitoefish Data
+
+```{r echo=FALSE, results='hide', message=FALSE}
+## Bonito data
+mqf <- read.csv("Mosquitoefish.csv")
+mqf2 <- filterD(mqf,Sitenum==2)
+mqf4 <- filterD(mqf,Sitenum==4)
+mqf9 <- filterD(mqf,Sitenum==9)
+```
+
+```{r echo=FALSE, results='hide', message=FALSE}
+## Mosquitofish Site 2
+
+################################################################################
+## See if results match that from Emili (Camona-Catot et al. (2014))
+lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
+uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=Inf,ts=1)
+svmqf2 <- list(Linf=36,K=2,t0=0,C=1.6,ts=0.9)
+fitmqf2 <- nls(SL~vbSO(Age2,Linf,K,t0,C,ts),data=mqf2,
+               start=svmqf2,lower=lwrbnd,upper=uprbnd,
+               algorithm="port",control=list(maxiter=100))
+## Linf=35.85, K=2.012, t0=-0.02, C=1.95, and ts=-0.118 ...
+## they all match (within rounding) except ts but it is off by +1
+cfmqf2 <- cbind(Est=coef(fitmqf2),confint(fitmqf2))
+################################################################################
+
+## Fit new model
+lwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
+uprbnd <- c(Linf=Inf,Kpr=Inf,t0=Inf,ts=1,NGT=1)
+svmqf2p <- list(Linf=36,Kpr=2,t0=0,ts=0.9,NGT=0.5)
+fitmqf2p <- nls(SL~VBSCGF(Age2,Linf,Kpr,t0,ts,NGT),data=mqf2,
+             start=svmqf2p,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+cfmqf2p <- cbind(Est=coef(fitmqf2p),confint(fitmqf2p))
+```
+
+```{r echo=FALSE, results='hide', message=FALSE}
+## Mosquitofish Site 4
+
+################################################################################
+## See if results match that from Emili (Camona-Catot et al. (2014))
+lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
+uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=Inf,ts=1)
+svmqf4 <- list(Linf=36,K=2,t0=0,C=1.6,ts=0.9)
+fitmqf4 <- nls(SL~vbSO(Age2,Linf,K,t0,C,ts),data=mqf4,
+               start=svmqf4,lower=lwrbnd,upper=uprbnd,
+               algorithm="port",control=list(maxiter=100))
+## Linf=45.97, K=1.046, t0=-0.197, C=1.28, and ts=0.751 ... They do
+cfmqf4 <- cbind(Est=coef(fitmqf4),confint(fitmqf4))
+################################################################################
+
+## Fit new model to Mosquitofish Site 5 data
+lwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
+uprbnd <- c(Linf=Inf,Kpr=Inf,t0=Inf,ts=1,NGT=1)
+svmqf4p <- list(Linf=36,Kpr=2,t0=0,ts=0.9,NGT=0.5)
+fitmqf4p <- nls(SL~VBSCGF(Age2,Linf,Kpr,t0,ts,NGT),data=mqf4,
+             start=svmqf4p,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+cfmqf4p <- cbind(Est=coef(fitmqf4p),confint(fitmqf4p))
+```
+
+```{r echo=FALSE, results='hide', message=FALSE}
+################################################################################
+## Purely Demonstration -- Mosquitofish Site 9
+## See if results match that from Emili (Camona-Catot et al. (2014))
+lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
+uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=Inf,ts=1)
+svmqf9 <- list(Linf=36,K=2,t0=0,C=1.6,ts=0.9)
+fitmqf9 <- nls(SL~vbSO(Age2,Linf,K,t0,C,ts),data=mqf9,
+               start=svmqf9,lower=lwrbnd,upper=uprbnd,
+               algorithm="port",control=list(maxiter=100))
+## Linf=41.54, K=1.314, t0=-0.214, C=0.621, and ts=-0.717 ... They do
+cfmqf9 <- cbind(Est=coef(fitmqf9),confint(fitmqf9))
+################################################################################
+
+## Fit new model to Mosquitofish Site 9 data
+lwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
+uprbnd <- c(Linf=Inf,Kpr=Inf,t0=Inf,ts=1,NGT=1)
+svmqf9p <- list(Linf=42,Kpr=1.6,t0=0,ts=0.7,NGT=0.05)
+fitmqf9p <- nls(SL~VBSCGF(Age2,Linf,Kpr,t0,ts,NGT),data=mqf9,
+             start=svmqf9p,lower=lwrbnd,upper=uprbnd,
+             algorithm="port",control=list(maxiter=100))
+cfmqf9p <- cbind(Est=coef(fitmqf9p),confint(fitmqf9p))
 ```

### 2016-05-22 code/SeasonalGrowth_Analysis.R

@@ -1,0 +1,237 @@
+################################################################################
+################################################################################
+##
+## Analysis script for ...
+##
+################################################################################
+################################################################################
+
+################################################################################
+## SETUP
+################################################################################
+## Load the R implementation of the Pauly et al. (1992) function
+source("code/iCalc_tpr.R")
+source("code/VBSCGF.R")
+## Create a Somers function
+vbSO <- function(t,Linf,K,t0,C,ts) {
+  if (length(Linf)==5) { K <- Linf[[2]]; t0 <- Linf[[3]]
+  C <- Linf[[4]]; ts <- Linf[[5]]; Linf <- Linf[[1]] }
+  St <- (C*K)/(2*pi)*sin(2*pi*(t-ts))
+  Sto <- (C*K)/(2*pi)*sin(2*pi*(t0-ts))
+  Linf*(1-exp(-K*(t-t0)-St+Sto))
+}
+
+
+################################################################################
+## Create the demo figure of the Somers model
+################################################################################
+pdf("results/Figure_1.PDF",width=4,height=4)
+par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
+ts <- 0.05; Linf <- 30; K <- 0.3; t0 <- -0.1
+C <- c(0,0.5,1,2)
+clr <- col2rgbt(rep("black",4),1/(1:4))
+t <- seq(0,3,length.out=299)
+plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],ts))~t,type="l",lwd=2,col=clr[1],
+     ylim=c(0,20),ylab="Mean Length",xlab="Age (years)",xaxt="n")
+axis(1,0:3)
[…]

### 2016-05-22 zzzOld/data/fromEmili/script Somers.R

@@ -1,0 +1,54 @@
+library(FSA);library(NCStats);library(nlstools);library(nls2)
+library(car);library(qpcR)
+
+setwd("D:/Emili2/R3 projectes/projectes MEC General/MEC 2006/Gambúsia latitud 2007/creix estacional/resultats diferenciant cohorts")
+file9<-read.csv("fig3.csv",sep = ";")
+
+file9$Females<-factor(file9$Females)
+file9$Sitenum<-factor(file9$Sitenum)
+sapply(file9, data.class)
+names(file9)
+summary(file9)
+head(file9)
+
+scatterplot(SL~ Age2  | Site, data=file9,smooth=T,reg.line=F,boxplots=F)
+
+####################    VBGF   #########################
+
+
+table(file9$Site,file9$Sitenum)
+
+# Altea Bourdigou Ebro Fluvià Millars Orb Segura Ter Vistre Xuquer
+# Select only 1; they are ordered by number/latitude
+SELECTION="Vistre"
+SELECTION="Orb"
+SELECTION="Bourdigou"
+SELECTION="Fluvià"
+SELECTION="Ter"
+SELECTION="Ebro"
+SELECTION="Millars"
+SELECTION="Xúquer"
+SELECTION="Altea"
+SELECTION="Segura"
+
+file8=file9[file9$Site==SELECTION, ]
+
+file8$len=file8$SL;file8$age=file8$Age2
+summary(file8$len);summary(file8$age)
+vbso <- vbFuns("Somers")
+
+scatterplot(SL~ Age2, data=file8,smooth=T,reg.line=F,boxplots=F)
+
+start <- data.frame(Linf = c(20, 50), K  = c(0, 1), t0 = c(-1, +1), C = c(0, 2), ts = c(0, +1))
+vbso1 <- nls2(len ~ vbso(age, Linf, K, t0, C, ts), data = file8, start = start, trace=T)
+summary(vbso1, correlation = TRUE)
+
+# confint(vbso1)
+Rsq(vbso1)
+ages <- seq(0,3,0.01) # choose age range
+plen <- predict(vbso1,data.frame(age=ages)) # predict length @ age
+plot(len~age,data=file8,pch=19,xlab="Age (years)",ylab="Standard Length (mm)",main=SELECTION)
+lines(ages,plen,lwd=2,col="red")
+
+remove(vbso1)
+

### 2016-05-25 code/SeasonalGrowth_Analysis.R
Started to format for Fisheries Research (v0.4)

@@ -9,18 +9,16 @@
 ################################################################################
 ## SETUP
 ################################################################################
+## Load the FSA package
+library(FSA)
+library(nlstools)
 ## Load the R implementation of the Pauly et al. (1992) function
 source("code/iCalc_tpr.R")
 source("code/VBSCGF.R")
 ## Create a Somers function
-vbSO <- function(t,Linf,K,t0,C,ts) {
-  if (length(Linf)==5) { K <- Linf[[2]]; t0 <- Linf[[3]]
-  C <- Linf[[4]]; ts <- Linf[[5]]; Linf <- Linf[[1]] }
-  St <- (C*K)/(2*pi)*sin(2*pi*(t-ts))
-  Sto <- (C*K)/(2*pi)*sin(2*pi*(t0-ts))
-  Linf*(1-exp(-K*(t-t0)-St+Sto))
-}
-
+vbSO <- vbFuns("Somers")
+## Set the random number seed so that the bootstraps stay reproducible
+set.seed(730987)
 
 ################################################################################
 ## Create the demo figure of the Somers model
@@ -91,7 +89,8 @@ SsvBon <- list(Linf=72,K=0.27,t0=0,C=0.6,ts=0.2)
 SfitBon <- nls(FL~vbSO(Age,Linf,K,t0,C,ts),data=bon,
                start=SsvBon,lower=Slwrbnd,upper=Suprbnd,
                algorithm="port",control=list(maxiter=100))
-ScfBon <- cbind(Est=coef(SfitBon),confint(SfitBon))
+SbootBon <- nlsBoot(SfitBon)
+ScfBon <- cbind(Est=coef(SfitBon),confint(SbootBon))
 
 ## 2. Fit new Pauly et al. (1992) function
 Plwrbnd <- c(Linf=0,Kpr=0,t0=-Inf,ts=0,NGT=0)
@@ -100,7 +99,8 @@ PsvBon <- list(Linf=72,Kpr=0.27,t0=0,ts=0.5,NGT=0.2)
[…]
